---
title: "TCGA COAD – Method Comparison: MatrixDecomp vs Lasso vs PCA/FA + OLS"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---

```{r setup, message=FALSE, warning=FALSE}
library(tidyverse)
library(glmnet)
library(psych)
library(factoextra)
library(knitr)
library(kableExtra)

opts_chunk$set(message = FALSE, warning = FALSE, fig.width = 8, fig.height = 5)

PROJECT_ROOT <- "~/multiomic_integration"
DATA_DIR     <- file.path(PROJECT_ROOT, "data/tcga")
EXPORT_DIR   <- file.path(DATA_DIR, "processed/kirc")
MD_RESULT_DIR <- file.path(PROJECT_ROOT, "sim/results/matrixdecomp/tcga")

dir.create(EXPORT_DIR, showWarnings = FALSE, recursive = TRUE)
```

## Data Loading

```{r load-data}
TMB_F1CDX <- readRDS(file.path(DATA_DIR, "TMB_F1CDX_data.rds"))
subtype_raw  <- TMB_F1CDX$kirc_tcga_pub

y_raw      <- subtype_raw$TMB_NONSYNONYMOUS
X_raw      <- subtype_raw |> select(-sampleId, -TMB_NONSYNONYMOUS) |> as.matrix()
sample_ids <- subtype_raw$sampleId

cat(sprintf("n = %d samples,  p = %d gene features\n", nrow(X_raw), ncol(X_raw)))
```

```{r}
TMB_F1CDX$kirc_tcga_pub |> filter(TMB_NONSYNONYMOUS > 200)
```


```{r tmb-distribution}
ggplot(tibble(TMB = y_raw), aes(x = TMB)) +
  geom_histogram(bins = 40, fill = "steelblue", colour = "white") +
  labs(title = "Distribution of TMB_NONSYNONYMOUS (COAD)",
       x = "TMB (mutations / Mb)", y = "Count") +
  theme_minimal()
```


## Preprocessing

```{r}
y_raw <- log(1 + y_raw)
```


```{r tmb-distribution}
ggplot(tibble(TMB = y_raw), aes(x = TMB)) +
  geom_histogram(bins = 40, fill = "steelblue", colour = "white") +
  labs(title = "Distribution of TMB_NONSYNONYMOUS (COAD)",
       x = "TMB (mutations / Mb)", y = "Count") +
  theme_minimal()
```


```{r preprocess}
set.seed(42)
n         <- nrow(X_raw)
train_idx <- sample(seq_len(n), size = floor(0.8 * n))
test_idx  <- setdiff(seq_len(n), train_idx)

X_train_raw <- X_raw[train_idx, ]
X_test_raw  <- X_raw[test_idx, ]
y_train_raw <- y_raw[train_idx]
y_test_raw  <- y_raw[test_idx]

# Column-wise standardisation using training statistics only
col_means <- colMeans(X_train_raw)
col_sds   <- apply(X_train_raw, 2, sd)

nonzero_var <- col_sds > 0
X_train_raw <- X_train_raw[, nonzero_var]
X_test_raw  <- X_test_raw[, nonzero_var]
col_means   <- col_means[nonzero_var]
col_sds     <- col_sds[nonzero_var]

X_train <- scale(X_train_raw, center = col_means, scale = col_sds)
X_test  <- scale(X_test_raw,  center = col_means, scale = col_sds)

# Standardise outcome (using training statistics)
y_mean    <- mean(y_train_raw)
y_sd      <- sd(y_train_raw)
y_train_z <- (y_train_raw - y_mean) / y_sd
y_test_z  <- (y_test_raw  - y_mean) / y_sd

cat(sprintf("Train: %d | Test: %d | Features after filtering: %d\n",
            length(train_idx), length(test_idx), sum(nonzero_var)))
```


```{r tmb-distribution}
ggplot(tibble(TMB = y_train_z), aes(x = TMB)) +
  geom_histogram(bins = 40, fill = "steelblue", colour = "white") +
  labs(title = "Distribution of TMB_NONSYNONYMOUS (COAD)",
       x = "TMB (mutations / Mb)", y = "Count") +
  theme_minimal()
```

## Export Data for Python (MatrixDecomp)

These CSVs are read by `python/tcga_matrixdecomp.py`.  Both matrices already have
zero-mean, unit-SD columns (training statistics applied to test).  The `y` column
is the standardised outcome.

```{r export}
train_df <- as.data.frame(X_train); train_df$y <- y_train_z
test_df  <- as.data.frame(X_test);  test_df$y  <- y_test_z

write.csv(train_df, file.path(EXPORT_DIR, "train.csv"), row.names = FALSE)
write.csv(test_df,  file.path(EXPORT_DIR, "test.csv"),  row.names = FALSE)

cat("Exported to", EXPORT_DIR, "\n")
cat("Run `python python/tcga_matrixdecomp.py` to generate MatrixDecomp results.\n")
```

## Helper Functions

```{r helpers}
# Returns the five metrics matching eval_performance_single_view() in Python.
#   structure_rse     – ||X_train - X_hat||²_F / ||X_train||²_F  (NA if no decomposition)
#   structure_95p_cov – 95% CI coverage of X_train reconstruction (NA if deterministic)
#   cov_norm_frob     – ||Cov_est - Cov_sample||²_F / p²  (NA if no decomposition)
#   outcome_mse       – mean squared error on test set
#   outcome_95p_cov   – 95% PI coverage on test set (NA if no intervals)
compute_metrics <- function(y_true, y_pred, method_name,
                            y_pred_lower    = NULL, y_pred_upper = NULL,
                            structure_rse     = NA_real_,
                            structure_95p_cov = NA_real_,
                            cov_norm_frob     = NA_real_) {
  outcome_mse     <- mean((y_true - y_pred)^2)
  outcome_95p_cov <- if (!is.null(y_pred_lower) && !is.null(y_pred_upper))
    mean(y_true >= y_pred_lower & y_true <= y_pred_upper)
  else
    NA_real_

  tibble(method            = method_name,
         structure_rse     = structure_rse,
         structure_95p_cov = structure_95p_cov,
         cov_norm_frob     = cov_norm_frob,
         outcome_mse       = outcome_mse,
         outcome_95p_cov   = outcome_95p_cov)
}

sample_cov <- crossprod(X_train) / (nrow(X_train) - 1)  # used by PCA and FA below

# Null model: predict training mean (= 0 on standardised scale).
metrics_null <- compute_metrics(y_test_z, rep(0, length(y_test_z)), "Null (mean)")
```

## Method 1: Lasso

```{r lasso, fig.height=4}
set.seed(42)
lasso_cv <- cv.glmnet(X_train, y_train_z, alpha = 1, nfolds = 10)
plot(lasso_cv, main = "Lasso CV")

pred_lasso_min <- as.numeric(predict(lasso_cv, newx = X_test, s = "lambda.min"))
pred_lasso_1se <- as.numeric(predict(lasso_cv, newx = X_test, s = "lambda.1se"))

n_coef_min <- sum(coef(lasso_cv, s = "lambda.min")[-1] != 0)
n_coef_1se <- sum(coef(lasso_cv, s = "lambda.1se")[-1] != 0)
cat(sprintf("Non-zero coefficients: lambda.min = %d | lambda.1se = %d\n",
            n_coef_min, n_coef_1se))

metrics_lasso <- bind_rows(
  compute_metrics(y_test_z, pred_lasso_min, sprintf("Lasso λ.min (%d coef)", n_coef_min)),
  compute_metrics(y_test_z, pred_lasso_1se, sprintf("Lasso λ.1se (%d coef)", n_coef_1se))
)

print(metrics_lasso)
```

## Method 2: PCA + OLS

PCA is fit on the (already standardised) training features; test observations are
projected using the same rotation.

```{r pca-ols}
pca_fit <- prcomp(X_train, center = FALSE, scale. = FALSE)

# Scree plot: variance explained
fviz_eig(pca_fit, ncp = 30, col.var = "steelblue",
         main = "PCA scree plot – COAD gene expression")

k_pca_values <- c(5, 10, 20, 50)
p <- ncol(X_train)

metrics_pca <- map_dfr(k_pca_values, function(k) {
  Lambda    <- pca_fit$rotation[, seq_len(k), drop = FALSE]
  sdev_k    <- pca_fit$sdev[seq_len(k)]
  Z_train   <- pca_fit$x[, seq_len(k), drop = FALSE]
  Z_test_df <- as.data.frame(predict(pca_fit, newdata = X_test)[, seq_len(k), drop = FALSE])
  Z_train_df <- as.data.frame(Z_train)

  # structure_rse: training reconstruction error
  X_hat         <- Z_train %*% t(Lambda)
  str_rse       <- sum((X_train - X_hat)^2) / sum(X_train^2)

  # cov_norm_frob: rank-k covariance approximation vs sample covariance
  cov_est       <- Lambda %*% diag(sdev_k^2, k) %*% t(Lambda)
  cov_frob      <- sum((cov_est - sample_cov)^2) / p^2

  # outcome: lm with prediction intervals on test set
  lm_fit    <- lm(y ~ ., data = cbind(y = y_train_z, Z_train_df))
  pred_int  <- predict(lm_fit, newdata = Z_test_df, interval = "prediction")

  compute_metrics(y_test_z, pred_int[, "fit"], sprintf("PCA+OLS (k=%d)", k),
                  y_pred_lower    = pred_int[, "lwr"],
                  y_pred_upper    = pred_int[, "upr"],
                  structure_rse   = str_rse,
                  cov_norm_frob   = cov_frob)
})

metrics_pca
```

## Method 3: Factor Analysis + OLS

Factor analysis is fit on the standardised training set; factor scores on new data
are obtained via `fa_fit$weights`.

```{r fa-ols}
k_fa_values <- c(5, 10, 20)

metrics_fa <- map_dfr(k_fa_values, function(k) {
  fa_fit     <- psych::fa(X_train, nfactors = k, rotate = "varimax",
                           scores = "regression", fm = "minres")
  Lambda     <- unclass(fa_fit$loadings)          # p × k
  Z_train_df <- as.data.frame(fa_fit$scores)
  Z_test_df  <- as.data.frame(X_test %*% fa_fit$weights)
  colnames(Z_test_df) <- colnames(Z_train_df)

  # structure_rse: training reconstruction error
  X_hat    <- fa_fit$scores %*% t(Lambda)
  str_rse  <- sum((X_train - X_hat)^2) / sum(X_train^2)

  # cov_norm_frob: Lambda @ Lambda' vs sample covariance (shared-factor part only)
  cov_est  <- Lambda %*% t(Lambda)
  cov_frob <- sum((cov_est - sample_cov)^2) / p^2

  # outcome: lm with prediction intervals on test set
  lm_fit   <- lm(y ~ ., data = cbind(y = y_train_z, Z_train_df))
  pred_int <- predict(lm_fit, newdata = Z_test_df, interval = "prediction")

  compute_metrics(y_test_z, pred_int[, "fit"], sprintf("FA+OLS (k=%d)", k),
                  y_pred_lower    = pred_int[, "lwr"],
                  y_pred_upper    = pred_int[, "upr"],
                  structure_rse   = str_rse,
                  cov_norm_frob   = cov_frob)
})

metrics_fa
```

## Load MatrixDecomp Results (Python)

MatrixDecomp (Bayesian supervised factor model) is trained in Python via
`python/tcga_matrixdecomp.py`, which reads the CSVs exported above.
Results are saved to `sim/results/matrixdecomp/tcga/`.

```{r load-matrixdecomp}
# Expects predictions written by data/tcga/tcga_matrixdecomp.qmd.
# structure_rse / structure_95p_cov / cov_norm_frob are not yet computed in the
# Python notebook and will be NA until added there.
md_metrics_path_sparse <- file.path(MD_RESULT_DIR, "metrics_20260418-233259.csv")
md_metrics_path_dense <- file.path(MD_RESULT_DIR, "metrics_20260422-111536.csv")
md_metrics_path_cv <- file.path(MD_RESULT_DIR, "metrics_20260422-143856.csv")

path_list <- c(md_metrics_path_sparse, md_metrics_path_dense, md_metrics_path_cv)
path_types <- c("_sparse", "_dense", "_cv")

matrix_decomp_results <- lapply(
  1:length(path_list),
  function(i) {
    md_metrics_path <- path_list[i]
    if (file.exists(md_metrics_path)) {
      md_metrics_raw <- read.csv(md_metrics_path)
      metrics_md <- bind_cols(tibble(
        method = paste0("MatrixDecomp", path_types[i])),
        md_metrics_raw)
      # k_val     <- md_metrics_raw$k[1]
      # pred_path <- file.path(MD_RESULT_DIR, sprintf("predictions_k%d.csv", k_val))
      # preds     <- read.csv(pred_path)
      # metrics_md <- compute_metrics(
      #   preds$y_true, preds$y_pred_mean,
      #   sprintf("MatrixDecomp (k=%d)", k_val),
      #   y_pred_lower = preds$`y_pred_q2.5`,
      #   y_pred_upper = preds$`y_pred_q97.5`
      # )
    } else {
      message("MatrixDecomp results not found in ", MD_RESULT_DIR,
              "\nRender data/tcga/tcga_matrixdecomp.qmd first.")
      metrics_md <- tibble(method = character(), structure_rse = double(),
                           structure_95p_cov = double(), cov_norm_frob = double(),
                           outcome_mse = double(), outcome_95p_cov = double())
    }
  }
) |> 
  bind_rows()
```

## Comparison

```{r comparison-table}
all_metrics <- bind_rows(
  metrics_null,
  metrics_lasso,
  metrics_pca,
  metrics_fa,
  matrix_decomp_results
  # metrics_md
) |>
  arrange(outcome_mse)

all_metrics |>
  mutate(across(where(is.double), \(x) round(x, 4))) |>
  kbl(caption = "Test-set prediction metrics (standardised outcome scale)") |>
  kable_styling(bootstrap_options = c("striped", "hover", "condensed")) |>
  row_spec(which.min(all_metrics$outcome_mse), bold = TRUE, background = "#d4edda")
```

```{r outcome-mse-plot, fig.height=6}
all_metrics |>
  mutate(method = fct_reorder(method, outcome_mse)) |>
  ggplot(aes(x = method, y = outcome_mse,
             fill = str_extract(method, "^[^\\s(]+"))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "TCGA COAD: TMB prediction – outcome MSE by method",
       x = NULL, y = "Test outcome MSE (standardised scale)") +
  theme_minimal(base_size = 12)
```

```{r outcome-coverage-plot, fig.height=6}
all_metrics |>
  filter(!is.na(outcome_95p_cov)) |>
  mutate(method = fct_reorder(method, outcome_95p_cov)) |>
  ggplot(aes(x = method, y = outcome_95p_cov,
             fill = str_extract(method, "^[^\\s(]+"))) +
  geom_col(show.legend = FALSE) +
  geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "TCGA COAD: TMB prediction – 95% PI coverage by method",
       subtitle = "Dashed line = nominal 0.95",
       x = NULL, y = "95% prediction interval coverage") +
  theme_minimal(base_size = 12)
```

```{r structure-rse-plot, fig.height=6}
all_metrics |>
  filter(!is.na(structure_rse)) |>
  mutate(method = fct_reorder(method, structure_rse)) |>
  ggplot(aes(x = method, y = structure_rse,
             fill = str_extract(method, "^[^\\s(]+"))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "TCGA COAD: training data reconstruction RSE by method",
       x = NULL, y = "Structure RSE") +
  theme_minimal(base_size = 12)
```

```{r cov-frob-plot, fig.height=6}
all_metrics |>
  filter(!is.na(cov_norm_frob)) |>
  mutate(method = fct_reorder(method, cov_norm_frob)) |>
  ggplot(aes(x = method, y = cov_norm_frob,
             fill = str_extract(method, "^[^\\s(]+"))) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  scale_fill_brewer(palette = "Set2") +
  labs(title = "TCGA COAD: covariance Frobenius error vs sample covariance",
       x = NULL, y = "||Cov_est - Cov_sample||²_F / p²") +
  theme_minimal(base_size = 12)
```

## Notes on Metrics

| Metric | Definition for real data | Methods |
|--------|--------------------------|---------|
| `structure_rse` | \|\|X_train − Ẑ Λ'\|\|²_F / \|\|X_train\|\|²_F | PCA, FA |
| `structure_95p_cov` | 95% PI coverage of X_train reconstruction | NA (deterministic methods) |
| `cov_norm_frob` | \|\|Λ Λ' − S_train\|\|²_F / p² where S_train = sample cov | PCA, FA |
| `outcome_mse` | Mean squared error on test set | All |
| `outcome_95p_cov` | 95% PI/CI coverage on test set | PCA+OLS, FA+OLS, MatrixDecomp |

Lasso has no explicit factor decomposition so structure and covariance metrics are NA.
MatrixDecomp structure metrics require additional outputs from `tcga_matrixdecomp.qmd`
(not yet computed).
```
