---
title: "Modeling"
author: "Jonathan Hori"
date: "2025-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# From GitHub - requires devtools
# devtools::install_github("niccoloanceschi/jafar")


library(jafar)
library(Rfast)
library(ggplot2)
library(pROC)
library(tidyverse)
library(latex2exp)

source("../R/data_methods.R")

jafar_data <- Data

output_dir = "./results/"
dir.create(output_dir) #, showWarnings = FALSE)
# getwd()
```

MCMC params
```{r}
tMCMC = 10000
tBurnIn = 5000
tThin = 10

mcmc_seed=1234
```



# JAFAR

## Data preprocessing

```{r}
getwd()
```

```{r}
# sample_ix <- sample(1:500, size = 100)

# sim_data <- readRDS("./data/sim_data_n100_snr1.1.rds")
# sim_data <- readRDS("./data/sim_data_n100_snr1.1_ywithview.rds")
# sim_data <- readRDS("./data/sim_data_n100_snr2.2_ywithview.rds")
# sim_data <- readRDS("./data/sim_data_n100_snr2.2_ywithview_sparseloadings.rds") 
# sim_data <- readRDS("./data/sim_data_n100p100_snr2.2_ywithview.rds") 
# sim_data <- readRDS("./data/sim_data_n1000p50_snr2.2_ywithview.rds") 
# sim_data <- readRDS("./data/n500p100_snr1.1/sim_data_ywithview_rep1.rds")
# sim_data <- readRDS("./data3/n500p100_snr2.2/sim_data_ywithview_rep1.rds")
sim_data <- readRDS("./data3/n500p100_snr2.2_sparse0.25/sim_data_ywithview_rep1.rds")

X_l <- sim_data$X_l
# readRDS("./data/sim_data_x.rds") |> 
  # lapply(function(x) x[sample_ix, ])
y <- sim_data$y
  # eadRDS("./data/sim_data_y.rds")[sample_ix]
```


```{r}
# X_l[[1]]
```


```{r}
X_l.clean <- standardize_views(X_l)
y.clean <- standardize_outcome(y)
```

```{r}
X_l.clean[[1]]
```


```{r}
X_l <- lapply(X_l,
              function(data) {
                colnames(data) <- seq_len(ncol(data))
                data
              })
colnames(X_l[[1]]) #<- seq_len(ncol(X_l[[1]]))
X_l.preprocess <- preprocess_X(X_l)
colnames(Data$X_m[[1]])
seq_len(ncol(X_l[[1]]))
ncol(X_l[[1]])

X_l[[1]]

lapply(1:3,
       function(i) {
         norm(X_l.preprocess$X_m[[i]] - X_l.clean[[i]], type = "F")
       })
norm(X_l.preprocess$X_m[[1]] - X_l.clean[[1]], type = "F")

X_l.preprocess$X_m[[1]]
X_l.clean[[1]]
```


## Single run

```{r}
system.time({
mcmc_supervised <- gibbs_jafar(X_m = X_l.clean, 
                               y = y.clean, 
                               # K0=K0, 
                               # K0_m=c(10, 10, 10),
                               tMCMC=tMCMC, 
                               tBurnIn=tBurnIn, 
                               tThin=tThin,
                               hyperparams = list(seed = mcmc_seed))
})
```

For 
**
N = 100
tMCMC = 10000
tBurnIn = 5000
tThin = 10

   user  system elapsed 
699.293 104.625 415.154 

**



### MatchAlign

```{r}
plot_loadings(mcmc_supervised,
              out_path = output_dir,
              out_shared = paste0(format(Sys.time(), "%m%d%y_%H%M"), "_shared_loadings"), 
              out_specific = paste0(format(Sys.time(), "%m%d%y_%H%M"), "_specific_loadings"))
```


```{r}
# mcmc_supervised <- multiviewMatchAlign(mcmc_supervised)
```


## Evaluate performance

Load individual model to check
```{r}
mcmc_supervised <- readRDS(paste0(
  "~/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0.25/models/", 
  "n500p100_snr2.2_sparse0.25_rep1.rds"))
```



### Structure reconstruction
Check reconstruction of shared and view-specific structure, which should be uniquely identified, unlike individual loading/score matrices.


```{r}
# jafar_posterior_means <- calc_jafar_posterior_summ(mcmc_supervised)
jafar_posterior_summaries <- summarise_post_samples(mcmc_supervised)
jafar_posterior_means <- jafar_posterior_summaries$mean
```


Compute latent structures

```{r}
X_l_means <- lapply(X_l, function(X) {
  means <- X |> colMeans()
  t(replicate(nrow(X), means))
  }
)

X_l_sds <- lapply(X_l, function(X) {
  sds <- X |> colVars() |> sqrt()
  diag(1 / sds)
  }
)
```


```{r}
# lapply(1:5,
#        function(x)
#          1/x)
```


```{r}
est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
                               jafar_posterior_means$view_scores,
                               jafar_posterior_means$shared_loadings,
                               jafar_posterior_means$view_loadings)

sim_obj <- calc_all_structures(sim_data$Z,
                               sim_data$Phi,
                               sim_data$Lambda_l,
                               sim_data$Gamma_l,
                               data_mean = X_l_means,
                               data_sd = X_l_sds)
```


```{r}
# lapply(X_l_means,
#        function(mean) {
#          t(replicate(10, mean))
#        })
# 
# X_l_sds[[1]] #|> diag()
# 
# is.matrix(est_obj[["covariances"]][[1]][[2]])
# 
# 
# est_obj[["data_reconstruction"]]
# sim_obj
# struct = sim_obj
# sim_obj[["data_reconstruction"]] <- lapply(1:length(struct[["data_reconstruction"]]),
#        function(l) {
#          (struct[["data_reconstruction"]][[l]] - X_l_means[[l]]) %*% X_l_sds[[l]]
#        })
# 
# sim_obj[["data_reconstruction"]]


```


#### Eval

Relative squared error in structures:

$$
\lVert\hat S_{shared} - S_{shared}\rVert^2 = \lVert\hat Z\hat\Lambda^{(l)T} - Z\Lambda^{(l)T}\rVert^2 \\
\frac{\lVert\hat S - S\rVert^2}{\lVert S\rVert^2}
$$

```{r}
jafar_structure.shared[[l]] |> 
  as_tibble() |> 
  mutate(n = row_number()) |> 
  pivot_longer(c(everything(), -n)) |> 
  ggplot() +
  geom_tile(aes(y = n, x = name, fill = value)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white")

sim_structure.shared[[l]] |> 
  as_tibble() |> 
  mutate(n = row_number()) |> 
  pivot_longer(c(everything(), -n)) |> 
  ggplot() +
  geom_tile(aes(y = n, x = name, fill = value)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white")



(jafar_structure.shared[[l]] - 
                sim_structure.shared[[l]]) |> summary() |> list()


for (l in 1:3) {
print((jafar_structure.shared[[l]] - 
                sim_structure.shared[[l]]) |> 
  as_tibble() |> 
  mutate(n = row_number()) |> 
  pivot_longer(c(everything(), -n)) |> 
  ggplot() +
  geom_tile(aes(y = n, x = name, fill = value)) +
  scale_fill_gradient2(low = "red", high = "blue", mid = "white") 
)
    
}
```


Difference in covariance

$$
\frac{\lVert\hat \Lambda^{(l)}\hat \Lambda^{(l)T} - \Lambda^{(l)}\Lambda^{(l)T} \rVert^2 }{ p_l}
$$

```{r}
eval_model(est_obj,
           sim_obj,
           "covariance",
           "difference_norm")
```


```{r}
plot_correlations(mcmc_supervised, X_l, out_path=output_dir, 
                  out_name = "correlations_n500p100_snr1.1_rep1")
```


### Data reconstruction

$$
\frac{\lVert\hat X^{(l)} - X^{(l)}\rVert^2}{\lVert X^{(l)}\rVert^2} \\
\lVert\hat X^{(l)} - X^{(l)}\rVert^2
$$

```{r}
# Sanity check: Verify simulated data structures reconstruct data
print("Relative squared error")
lapply(1:sim_data$L,
       function(l) {
         norm(sim_obj$data_reconstruction[[l]] -
                X_l[[l]], type = "F") ^ 2 /
           norm(X_l[[l]], type = "F") ^2
       })


# sim_obj$data_reconstruction[[1]]
# 
# print("Difference norm")
# lapply(1:sim_data$L,
#        function(l) {
#          norm(sim_structure.reconstructed[[l]] - 
#                 X_l[[l]], type = "F") ^ 2
#        })

# 
# eval_model(est_obj,
#            sim_obj,
#            "data_reconstruction",
#            "rse")

```

```{r}
sim_obj$data_reconstruction[[1]]

diag(1 / X_1_sds) %*% sim_obj$covariances[[1]][[1]] %*% diag(1 / X_1_sds)
diag(1 / X_1_sds)
est_obj$covariances[[1]]

sim_obj$covariances[[1]]

X_1_sds <- X_l[[1]] |> colVars() |> sqrt()


X_1_sds_sq <- X_l[[1]] |> colVars() 




X_1_sds_sq
X_l[[1]] |> colMeans()

X_l[[1]] |> scale(scale = X_1_sds)

sim_obj$joint_structure[[1]] |> scale(scale = X_1_sds)
est_obj$joint_structure[[1]]

norm(scale(sim_obj$joint_structure[[1]], scale = X_1_sds), type = "F")
norm(sim_obj$joint_structure[[1]], type = "F")

norm(scale(sim_obj$joint_structure[[1]], scale = X_1_sds) - est_obj$joint_structure[[1]], type = "F")^2 / norm(scale(sim_obj$joint_structure[[1]], scale = X_1_sds), type = "F")^2


diff <- diag(1 / X_1_sds) %*% sim_obj$covariances[[1]][[1]] %*% diag(1 / X_1_sds) - est_obj$covariances[[1]][[1]]
norm(diff, type = "F") / (ncol(diff) * nrow(diff))


scale(sim_obj$joint_structure[[1]], scale = X_1_sds) - est_obj$joint_structure[[1]]
```






```{r}
bind_rows(
  eval_model(est_obj,
           sim_obj,
           "joint_structure",
           "rse"),
  eval_model(est_obj,
           sim_obj,
           "view_structure",
           "rse"),
  eval_model(est_obj,
           sim_obj,
           "data_reconstruction",
           "rse"),
)
# ) |>
#   mutate(n = 1000,
#          p = 50,
#          snr_x = 2,
#          snr_y = 2) |>
#   print() |>
#   ggplot() +
#   geom_point(aes(x = n,
#                  y = result,
#                  group = quantity,
#                  color = quantity)) #|>
#   facet_grid(rows = vars())


```


### Outcome model coefficients

```{r}
# colMeans(mcmc_supervised$Theta)
# beta_sim
```



### Outcome prediction

```{r}
y_pred <- predict_y(X_l.clean, mcmc_supervised)

plot_predictions(y_pred, y.clean, mcmc_supervised,
                 out_path=output_dir, out_name="preds_sim4")
```

```{r}
# knitr::include_graphics(file.path(output_dir,'predictions.pdf'))
```

```{r}
plot_coefficients(y_pred)
```


Evaluate prediction performance

```{r}
y_pred.means <- colMeans(y_pred$mean)

```




```{r}
# y.clean
tibble(pred = y_pred.means,
       outcome = y.clean) |> 
  ggplot(aes(x = outcome, y = pred)) +
  geom_point() +
  stat_smooth(method = lm) + 
  geom_abline(slope=1, intercept = 0) +
  labs(x = "Simulated outcome",
       y = "Predicted outcome")
```



### Credible interval coverage

```{r}
# sim_obj$ > jafar_posterior_summaries$q2.5$shared_scores
```

```{r}
y_lower = apply(y_pred$mean, 2, quantile, probs = c(0.025))
y_upper = apply(y_pred$mean, 2, quantile, probs = c(0.975))

y_lower = apply(y_pred$mean, 2, quantile, probs = c(0.05))
y_upper = apply(y_pred$mean, 2, quantile, probs = c(0.95))
```


```{r}
eval_post_coverage <- function(data, 
                               lower, 
                               upper) {
  mean(data > lower & data < upper)
}
```

```{r}
eval_post_coverage(y.clean, y_lower, y_upper)
```


```{r}
tibble(pred = y_pred.means,
       outcome = y.clean,
       lower = y_lower,
       upper = y_upper) |> 
  arrange(pred) |> 
  mutate(id = row_number()) |> 
  ggplot() +
  geom_errorbar(aes(x = id, ymin = lower, ymax = upper)) +
  geom_point(aes(x = id, y = pred)) +
  geom_point(aes(x = id, y = outcome)) +
  # stat_smooth(method = lm) + 
  # geom_abline(slope=1, intercept = 0) +
  labs(x = "Simulated outcome",
       y = "Predicted outcome")
```


#-------------------------------
# JAFAR Example Dataset

```{r}
data("Data", package = "jafar")

M <- length(Data$X_m)
p_m <- sapply(Data$X_m,ncol)
n <- nrow(Data$X_m[[1]])
nTest <- nrow(Data$X_m_test[[1]])
```


```{r}
preprocessed_X <- preprocess_X(Data$X_m,X_m_test=Data$X_m_test,copula=F)
X_m <- preprocessed_X$X_m
X_m_test <- preprocessed_X$X_m_test

preprocessed_y <- preprocess_y(Data$yTrain,yTest=Data$yTest)

y <- preprocessed_y$yTrain
yTest <- preprocessed_y$yTest
```


```{r}
preprocessed_X$X_m[[1]] == X_m[[1]]
X_m[[1]] |> colVars()
```


```{r}
tMCMC = 10000
tBurnIn = 5000
tThin = 10

mcmc_seed=1234

K0 = 25
K0_m = rep(20,M)
```

```{r}
mcmc_supervised.example <- gibbs_jafar(X_m, y=y, K0=K0, K0_m=K0_m,
                             tMCMC=tMCMC, tBurnIn=tBurnIn, tThin=tThin,
                             hyperparams = list(seed = mcmc_seed))
```


```{r}
y_pred.jafar <- predict_y(X_m, mcmc_supervised.example)
y_pred.means.jafar <- colMeans(y_pred.jafar$mean)

tibble(pred = y_pred.means.jafar,
       outcome = y) |> 
  ggplot(aes(x = outcome, y = pred)) +
  geom_point() +
  stat_smooth(method = lm) + 
  geom_abline(slope=1, intercept = 0) +
  labs(x = "Simulated outcome",
       y = "Predicted outcome")
```

## Evaluate structure/data reconstruction, covariance matrix

```{r}
jafar_posterior_means.example <- calc_jafar_posterior_means(mcmc_supervised.example)

est_obj <- calc_all_structures(jafar_posterior_means.example$shared_scores,
                               jafar_posterior_means.example$view_scores,
                               jafar_posterior_means.example$shared_loadings,
                               jafar_posterior_means.example$view_loadings)

sim_obj <- calc_all_structures(Data$eta,
                               Data$phi_m,
                               Data$Lambda_m,
                               Data$Gamma_m)
```


```{r}
bind_rows(
  eval_model(est_obj,
           sim_obj,
           "joint_structure",
           "rse"),
  eval_model(est_obj,
           sim_obj,
           "view_structure",
           "rse"),
  eval_model(est_obj,
           sim_obj,
           "data_reconstruction",
           "rse"),
)

eval_model(est_obj,
           sim_obj,
           "covariance",
           "difference_norm")
```


```{r}
sim_obj$joint_structure[[1]] #|> scale(scale = X_1_sds)
est_obj$joint_structure[[1]]


sim_obj$data_reconstruction[[1]] #|> colMeans()#|> scale(scale = X_1_sds)
est_obj$data_reconstruction[[1]] #|> colMeans()


sim_obj$covariances[[1]][[1]]
est_obj$covariances[[1]][[1]]
```



```{r}
# output_dir = file.path(dirname(getSourceEditorContext()$path), 'results')
# dir.create(output_dir, showWarnings = FALSE)
plot_correlations(mcmc_supervised.example, X_m, out_path=output_dir)
```



## Interpretation
```{r}
mcmc_supervised.example <- multiviewMatchAlign(mcmc_supervised.example)
```


#-------------------------------
# Multiple datasets

Only check SNR = 2, ignore N = 1000

```{r}
n_array <- c(50, 100, 500)
# n_array <- c(50, 100, 500, 1000)
p_array <- c(50, 100, 1000)
snr_x_array <- c(2) #c(2, 1, 0.5)
snr_y_array <- c(2) #c(2, 1, 0.5)
# snr_x_array <- c(2, 1, 0.5)
# snr_y_array <- c(2, 1, 0.5)
reps = 10
loading_sparsity = 0

sim_grid <- expand.grid(n = n_array,
            p_l = p_array,
            snr_x = snr_x_array,
            snr_y = snr_y_array,
            rep = 1:reps)

data_name_base <- "sim_data_ywithview_rep%s.rds"
data_path_base <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                      "data3",
                      "n%sp%s_snr%s.%s_sparse%s")

out_path_base <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                           "results",
                           paste0("jafar/sparse", loading_sparsity))
out_name_base <- "n%sp%s_snr%s.%s_sparse%s_rep%s.rds"


log_dir <- "logs"

log_message <- function(...) {
  msg <- paste(Sys.time(), paste(..., collapse = " "), "\n")
  cat(msg, file = log_file, append = TRUE)
}
```


Raw code below. Attempting to parallelize below this.
```{r}
# eval_table.collected <- tibble()
# mcmc_objects <- list()
# 
# mapply(function(n, p, snr_x, snr_y) {
#   for (rep in 1:reps) {
#     # Load + process data
#     run_data <- load_data(n, p, snr_x, snr_y, rep)
#     X_l <- run_data$X_l
#     y <- run_data$y
#     X_l.clean <- standardize_views(X_l)
#     y.clean <- standardize_outcome(y)
#     
#     # use default K0. 
#     # FIT MODEL
#     tik <- Sys.time()
#     mcmc_supervised <- gibbs_jafar(X_m = X_l.clean, 
#                                    y = y.clean, 
#                                    # K0=K0, 
#                                    # K0_m=c(10, 10, 10),
#                                    tMCMC=tMCMC, 
#                                    tBurnIn=tBurnIn, 
#                                    tThin=tThin,
#                                    hyperparams = list(seed = mcmc_seed))
#     tok <- Sys.time()
#     runtime <- tok - tik
#     
#     # Average over MCMC samples
#     jafar_posterior_means <- calc_jafar_posterior_means(mcmc_supervised)
#     
#     # Calculate structures from fitted model and simulated data
#     est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
#                                    jafar_posterior_means$view_scores,
#                                    jafar_posterior_means$shared_loadings,
#                                    jafar_posterior_means$view_loadings)
#     
#     sim_obj <- calc_all_structures(run_data$Z,
#                                    run_data$Phi,
#                                    run_data$Lambda_l,
#                                    run_data$Gamma_l)
#     
#     eval_table <- bind_rows(
#       eval_model(est_obj,
#                  sim_obj,
#                  "joint_structure",
#                  "rse"),
#       eval_model(est_obj,
#                  sim_obj,
#                  "view_structure",
#                  "rse"),
#       eval_model(est_obj,
#                  sim_obj,
#                  "data_reconstruction",
#                  "rse"),
#     ) |> 
#       mutate(n = n,
#              p = p,
#              snr_x = snr_x,
#              snr_y = snr_y,
#              rep = rep,
#              time = as.numeric(difftime(tok, tik, units = "secs")))
#     
#     eval_table.collected <- bind_rows(eval_table.collected,
#                                       eval_table)
#     mcmc_objects[paste0("n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep)] <- mcmc_supervised
#   }
# },
# sim_grid$n,
# sim_grid$p_l,
# sim_grid$snr_x,
# sim_grid$snr_y)
```


```{r}

eval_performance <- function(mod, run_data) {
                             # n, p, snr_x, snr_y, rep, sparsity) {
  # X_l <- run_data$X_l
        # y <- run_data$y
        # log_message("Standardizing")
        X_l.clean <- standardize_views(run_data$X_l)
        
        # mod <- readRDS(model_out_data_path)
        
        # Calculate mean and sd of each feature - for rescaling simulated structures
        X_l_means <- lapply(run_data$X_l, function(X) {
          means <- X |> colMeans()
          t(replicate(nrow(X), means))
        }
        )
        
        X_l_sds_inv <- lapply(run_data$X_l, function(X) {
          sds <- X |> colVars() |> sqrt()
          diag(1 / sds)
        }
        )
        
        # Calculate posterior summaries for fitted model
        jafar_posterior_summaries <- summarise_post_samples(mod)
        
        # Calculate structures from fitted model and simulated data
        jafar_posterior_means <- jafar_posterior_summaries$mean
        est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
                                       jafar_posterior_means$view_scores,
                                       jafar_posterior_means$shared_loadings,
                                       jafar_posterior_means$view_loadings)
        sim_obj <- calc_all_structures(run_data$Z,
                                       run_data$Phi,
                                       run_data$Lambda_l,
                                       run_data$Gamma_l,
                                       data_mean = X_l_means,
                                       data_sd = X_l_sds_inv)
        
        # eval_table <- eval_model(est_obj,
        #                          sim_obj,
        #                          "covariance",
        #                          "difference_norm") |> 
        #   mutate(n = n,
        #          p = p,
        #          snr_x = snr_x,
        #          snr_y = snr_y,
        #          rep = rep
        #   )
        
        # Eval outcome metrics
        y.clean <- standardize_outcome(run_data$y)
        y_pred <- predict_y(X_l.clean, mod)
        y_lower = apply(y_pred$mean, 2, quantile, probs = c(0.025))
        y_upper = apply(y_pred$mean, 2, quantile, probs = c(0.975))
        
        outcome_coverage <- eval_post_coverage(y.clean, y_lower, y_upper)
        
        y_pred.means <- colMeans(y_pred$mean)
        y_mse <- mean((y_pred.means - y.clean)^2)
        
        # Aggregate metrics
        eval_table <- bind_rows(
          eval_model(est_obj,
                     sim_obj,
                     "joint_structure",
                     "rse"),
          eval_model(est_obj,
                     sim_obj,
                     "view_structure",
                     "rse"),
          eval_model(est_obj,
                     sim_obj,
                     "data_reconstruction",
                     "rse"),
          tibble_row(
            l = 0,
            quantity = "outcome",
            metric = "95p_coverage",
            result = outcome_coverage
          ),
          tibble_row(
            l = 0,
            quantity = "outcome",
            metric = "mse",
            result = y_mse
          )
        ) 
        # |> 
        #   mutate(n = n,
        #          p = p,
        #          snr_x = snr_x,
        #          snr_y = snr_y,
        #          rep = rep,
        #          sparsity = loading_sparsity
        #          # time = as.numeric(difftime(tok, tik, units = "secs"))
        #          )
        
        return(eval_table)
}


run_and_eval_model <- function(n, p, snr_x, snr_y, rep, sparsity,
                               in_path_base,
                               in_name_base,
                               out_path_base,
                               out_name_base) {
  model_out_data_path <- file.path(out_path_base,
                             "models",
                             sprintf(out_name_base,  n, p, snr_x, snr_y, sparsity, rep))
  # Skip dataset if model already run
  if (file.exists(model_out_data_path)) {NA}
  else {
  
  # n <- params$n
  # p <- params$p
  # snr_x <- params$snr_x
  # snr_y <- params$snr_y
  
  # Load + process data
  run_data <- load_data(in_path_base, in_name_base,
                        n, p, snr_x, snr_y, sparsity, rep)
  X_l <- run_data$X_l
  y <- run_data$y
  log_message("Standardizing")
  X_l.clean <- standardize_views(X_l)
  y.clean <- standardize_outcome(y)
  log_message("Done standardizing")
  
  # use default K0. 
  # FIT MODEL
  log_message("Fitting:")
  tik <- Sys.time()
  tryCatch({
  mcmc_supervised <- gibbs_jafar(X_m = X_l.clean, 
                                 y = y.clean, 
                                 # K0=K0, 
                                 # K0_m=c(10, 10, 10),
                                 tMCMC=tMCMC, 
                                 tBurnIn=tBurnIn, 
                                 tThin=tThin,
                                 hyperparams = list(seed = mcmc_seed))
  }, error = function(e) {
    message(e)
    log_message(e)
  })
  tok <- Sys.time()
  runtime <- tok - tik
  
  # Export model
  # model_out_data_path <- file.path(out_path_base,
  #                            "models",
  #                            sprintf(out_name_base,  n, p, snr_x, snr_y, rep))
  log_message("Writing model:", as.character(model_out_data_path))
  dir.create(dirname(model_out_data_path), recursive = TRUE, showWarnings = TRUE)
  saveRDS(mcmc_supervised,
          model_out_data_path)
  
  tryCatch({
    # log_message("Evaluating metrics")
    # Average over MCMC samples
    # log_message("→ calc_jafar_posterior_means")
    # jafar_posterior_means <- calc_jafar_posterior_means(mcmc_supervised)
    # # Calculate mean and sd of each feature - for rescaling simulated structures
    # X_l_means <- lapply(run_data$X_l, function(X) {
    #   means <- X |> colMeans()
    #   t(replicate(nrow(X), means))
    # }
    # )
    # 
    # X_l_sds_inv <- lapply(run_data$X_l, function(X) {
    #   sds <- X |> colVars() |> sqrt()
    #   diag(1 / sds)
    # }
    # )
    # 
    # # Calculate structures from fitted model and simulated data
    # # log_message("→ calc_all_structures est_obj")
    # est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
    #                                jafar_posterior_means$view_scores,
    #                                jafar_posterior_means$shared_loadings,
    #                                jafar_posterior_means$view_loadings)
    # # log_message("→ calc_all_structures sim_obj")
    # sim_obj <- calc_all_structures(run_data$Z,
    #                                run_data$Phi,
    #                                run_data$Lambda_l,
    #                                run_data$Gamma_l,
    #                                data_mean = X_l_means,
    #                                data_sd = X_l_sds_inv)
    # 
    # # log_message("→ eval_model + bind_rows")
    # eval_table <- bind_rows(
    #   eval_model(est_obj,
    #              sim_obj,
    #              "joint_structure",
    #              "rse"),
    #   eval_model(est_obj,
    #              sim_obj,
    #              "view_structure",
    #              "rse"),
    #   eval_model(est_obj,
    #              sim_obj,
    #              "data_reconstruction",
    #              "rse"),
    # ) |> 
    #   mutate(n = n,
    #          p = p,
    #          snr_x = snr_x,
    #          snr_y = snr_y,
    #          sparsity = sparsity,
    #          rep = rep,
    #          time = as.numeric(difftime(tok, tik, units = "secs")))
    
    eval_table <- eval_performance() |> 
          mutate(n = n,
                 p = p,
                 snr_x = snr_x,
                 snr_y = snr_y,
                 rep = rep,
                 sparsity = loading_sparsity,
                 time = as.numeric(difftime(tok, tik, units = "secs"))
                 )
    
    covariance_eval <- eval_model(est_obj,
                                  sim_obj,
                                  "covariance",
                                  "difference_norm") |> 
      mutate(n = n,
             p = p,
             snr_x = snr_x,
             snr_y = snr_y,
             sparsity = sparsity,
             rep = rep
      )
    
    
    ### Export results
    # eval_table.collected <- bind_rows(eval_table.collected,
    #                                   eval_table)
    metric_out_data_path <- file.path(out_path_base,
                                      "metrics",
                                      sprintf(out_name_base, n, p, snr_x, snr_y, sparsity, rep))
    cov_out_data_path <- file.path(out_path_base,
                                   "covariances",
                                   sprintf(out_name_base, n, p, snr_x, snr_y, sparsity, rep))
    log_message("Writing metrics:", as.character(metric_out_data_path))
    dir.create(dirname(metric_out_data_path), recursive = TRUE, showWarnings = TRUE)
    dir.create(dirname(cov_out_data_path), recursive = TRUE, showWarnings = TRUE)
    
    stopifnot(exists("eval_table"))
    
    saveRDS(eval_table,
            metric_out_data_path)
    
    saveRDS(covariance_eval,
            cov_out_data_path)
  }, error = function(e) {
    message(e)
    log_message(e)
  })
  }
  
}
```


```{r}

# test_run_and_eval_model <- function(n, p, snr_x, snr_y, rep,
#                                     in_path_base,
#                                     in_name_base,
#                                     out_path_base,
#                                     out_name_base) {
#   log_message("Starting run")
#   load_data(in_path_base, in_name_base,
#             n, p, snr_x, snr_y, rep)
#   log_message("data loaded")
#   
#   log_message(paste0(n, p, snr_x, snr_y, rep))
#   eval_table <- tibble(c(n, p, snr_x, snr_y, rep))
#   log_message(paste0(out_path_base, "mod_n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep))
#   
#   saveRDS(eval_table, 
#           paste0(out_path_base, "/mod_n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep, ".RDS"))
# }


```


### Fit JAFAR in parallel
```{r}
library(foreach)
library(doParallel)
```


```{r}

n_cores <- detectCores()
n_cores

cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)

timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
dir.create(log_dir, showWarnings = FALSE)
log_file <- file.path(log_dir, paste0("run_", timestamp, ".log"))
file.create(log_file)

# sim_grid_test <- sim_grid |> filter(snr_x == 2, #0.5, 
#                                     # (snr_y == 0.5 | snr_y == 1), 
#                                     n < 1000,
#                                     p_l < 10000)

tik <- Sys.time()
# Run sim
tryCatch({
  jafar_sim_results <- foreach(i = 1:nrow(sim_grid),
                               .combine = bind_rows,
                               .packages = c("tidyverse",
                                             "jafar",
                                             "Rfast")#,
                               # .export = c("load_data",
                               #             "standardize_views",
                               #             "calc_jafar_posterior_means",
                               #             "calc_all_structures",
                               #             "tMCMC",
                               #             "tBurnIn",
                               #             "tThin",
                               #             "mcmc_seed")
  ) %dopar% {
    log_message("Worker:", paste(Sys.info()[['nodename']], Sys.getpid(), sep='-'))
    params <- sim_grid[i, ]
    run_and_eval_model(params$n,
                       params$p,
                       params$snr_x,
                       params$snr_y,
                       params$rep,
                       loading_sparsity,
                       data_path_base,
                       data_name_base,
                       out_path_base,
                       out_name_base)
  }
  
  message("Stopping cluster")
  stopCluster(cl = cluster)
}, error = function(e){
  message(e)
  log_message("Worker:", paste(Sys.info()[['nodename']], Sys.getpid(), sep='-'), e)
  message("Stopping cluster")
  stopCluster(cl = cluster)
}
)
tok <- Sys.time()

print(tok - tik)
```


```{r}
jafar_sim_results
```



```{r}
for (n in n_array) {
  for (p in p_array) {
    for (rep in 1:reps) {
      # Load + process data
      run_data <- load_data(n, p, snr_x, snr_y)
      X_l <- run_data$X_l
      y <- run_data$y
      X_l.clean <- standardize_views(X_l)
      y.clean <- standardize_outcome(y)
      
      # use default K0
      mcmc_supervised <- gibbs_jafar(X_m = X_l.clean, 
                                     y = y.clean, 
                                     # K0=K0, 
                                     # K0_m=c(10, 10, 10),
                                     tMCMC=tMCMC, 
                                     tBurnIn=tBurnIn, 
                                     tThin=tThin,
                                     hyperparams = list(seed = mcmc_seed))
      
      # Average over MCMC samples
      jafar_posterior_means <- calc_jafar_posterior_means(mcmc_supervised)
      
      # Calculate structures from fitted model and simulated data
      est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
                                     jafar_posterior_means$view_scores,
                                     jafar_posterior_means$shared_loadings,
                                     jafar_posterior_means$view_loadings)
      
      sim_obj <- calc_all_structures(run_data$Z,
                                     run_data$Phi,
                                     run_data$Lambda_l,
                                     run_data$Gamma_l)
      
      eval_table <- bind_rows(
        eval_model(est_obj,
                   sim_obj,
                   "joint_structure",
                   "rse"),
        eval_model(est_obj,
                   sim_obj,
                   "view_structure",
                   "rse"),
        eval_model(est_obj,
                   sim_obj,
                   "data_reconstruction",
                   "rse"),
      ) |> 
        mutate(n = n,
               p = p,
               snr_x = snr_x,
               snr_y = snr_y,
               rep = rep)
      
      eval_table.collected <- bind_rows(eval_table.collected,
                                        eval_table)
      mcmc_objects[paste0("n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep)] <- mcmc_supervised
      
    }
  }
}
```


## Evaluate

Load all metric files and plot

```{r}
getwd()
####### Metrics2 - reevaluating
jafar_metric_filenames_sparse0 <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0/metrics3", 
                               pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0 <- lapply(jafar_metric_filenames_sparse0,
                            function(x) readRDS(x)) |> 
  bind_rows()


jafar_metric_filenames_sparse0.25 <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0.25/metrics2", 
                               pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0.25 <- lapply(jafar_metric_filenames_sparse0.25,
                            function(x) readRDS(x)) |> 
  bind_rows()
```

```{r}
metric_filenames[1]
```

```{r}
######## With runtimes
metric_filenames_sparse0_withtime <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0/metrics/", pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0_withtime <- lapply(metric_filenames_sparse0_withtime,
                            function(x) readRDS(x)) |> 
  bind_rows()


metric_filenames_sparse0.25_withtime <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0.25/metrics/", pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0.25_withtime <- lapply(metric_filenames_sparse0.25_withtime,
                            function(x) readRDS(x)) |> 
  bind_rows()
```



```{r}
getwd()
cov_filenames <- list.files("./results/jafar/covariances2/", pattern = "\\.rds$", full.names = TRUE)

jafar_cov_results <- lapply(cov_filenames,
                            function(x) readRDS(x)) |> 
  bind_rows()
```

```{r}
metric_filenames
```


### Evaluate full metrics for fitted models
```{r}
model_filenames <- list.files("./results/jafar/models/", pattern = "\\.rds$", full.names = TRUE)

# jafar_sim_models <- lapply(metric_filenames,
#                             function(x) readRDS(x)) |> 
#   bind_rows()
```

Only need to run below if metrics/covariances weren't calculated (properly) originally, or if some metrics need to be readded for fitted models.
```{r}
# i = 1
metrics_agg <- lapply(
  1:nrow(sim_grid),
  function(i) {
    # Parameters for given run
    params <- sim_grid[i, ]
    n <- params$n
    p <- params$p
    snr_x <- params$snr_x
    snr_y <- params$snr_y
    rep <- params$rep
    
    # Only evaluate parameter combinations with a fitted model
    model_out_data_path <- file.path(out_path_base,
                                     "models",
                                     sprintf(out_name_base,  n, p, snr_x, snr_y, loading_sparsity, rep))
    
    
    if (file.exists(model_out_data_path)) {
      # print(model_out_data_path)
      
      tryCatch({
        run_data <- load_data(data_path_base, data_name_base,
                              n, p, snr_x, snr_y, loading_sparsity, rep,
                              log = FALSE)
        mod <- readRDS(model_out_data_path)
        
        
        eval_table <- eval_performance(mod, run_data) |> 
          mutate(n = n,
                 p = p,
                 snr_x = snr_x,
                 snr_y = snr_y,
                 rep = rep,
                 sparsity = loading_sparsity
                 # time = as.numeric(difftime(tok, tik, units = "secs"))
                 )
          
        # # X_l <- run_data$X_l
        # # y <- run_data$y
        # # log_message("Standardizing")
        # X_l.clean <- standardize_views(run_data$X_l)
        # 
        # mod <- readRDS(model_out_data_path)
        # 
        # # Calculate mean and sd of each feature - for rescaling simulated structures
        # X_l_means <- lapply(run_data$X_l, function(X) {
        #   means <- X |> colMeans()
        #   t(replicate(nrow(X), means))
        # }
        # )
        # 
        # X_l_sds_inv <- lapply(run_data$X_l, function(X) {
        #   sds <- X |> colVars() |> sqrt()
        #   diag(1 / sds)
        # }
        # )
        # 
        # # Calculate posterior summaries for fitted model
        # jafar_posterior_summaries <- summarise_post_samples(mod)
        # 
        # # Calculate structures from fitted model and simulated data
        # jafar_posterior_means <- jafar_posterior_summaries$mean
        # est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
        #                                jafar_posterior_means$view_scores,
        #                                jafar_posterior_means$shared_loadings,
        #                                jafar_posterior_means$view_loadings)
        # sim_obj <- calc_all_structures(run_data$Z,
        #                                run_data$Phi,
        #                                run_data$Lambda_l,
        #                                run_data$Gamma_l,
        #                                data_mean = X_l_means,
        #                                data_sd = X_l_sds_inv)
        # 
        # # eval_table <- eval_model(est_obj,
        # #                          sim_obj,
        # #                          "covariance",
        # #                          "difference_norm") |> 
        # #   mutate(n = n,
        # #          p = p,
        # #          snr_x = snr_x,
        # #          snr_y = snr_y,
        # #          rep = rep
        # #   )
        # 
        # # Eval outcome metrics
        # y.clean <- standardize_outcome(run_data$y)
        # y_pred <- predict_y(X_l.clean, mod)
        # y_lower = apply(y_pred$mean, 2, quantile, probs = c(0.025))
        # y_upper = apply(y_pred$mean, 2, quantile, probs = c(0.975))
        # 
        # outcome_coverage <- eval_post_coverage(y.clean, y_lower, y_upper)
        # 
        # y_pred.means <- colMeans(y_pred$mean)
        # y_mse <- mean((y_pred.means - y.clean)^2)
        # 
        # # Aggregate metrics
        # eval_table <- bind_rows(
        #   eval_model(est_obj,
        #              sim_obj,
        #              "joint_structure",
        #              "rse"),
        #   eval_model(est_obj,
        #              sim_obj,
        #              "view_structure",
        #              "rse"),
        #   eval_model(est_obj,
        #              sim_obj,
        #              "data_reconstruction",
        #              "rse"),
        #   tibble_row(
        #     l = 0,
        #     quantity = "outcome",
        #     metric = "95p_coverage",
        #     result = outcome_coverage
        #   ),
        #   tibble_row(
        #     l = 0,
        #     quantity = "outcome",
        #     metric = "mse",
        #     result = y_mse
        #   )
        # ) |> 
        #   mutate(n = n,
        #          p = p,
        #          snr_x = snr_x,
        #          snr_y = snr_y,
        #          rep = rep,
        #          sparsity = loading_sparsity
        #          # time = as.numeric(difftime(tok, tik, units = "secs"))
        #          )
  
        # covariance_eval <- eval_model(est_obj,
        #                               sim_obj,
        #                               "covariance",
        #                               "difference_norm") |>
        #   mutate(n = n,
        #          p = p,
        #          snr_x = snr_x,
        #          snr_y = snr_y,
        #          rep = rep
        #   )
        
        
      
  
        ### Export results
        # eval_table.collected <- bind_rows(eval_table.collected,
        #                                   eval_table)
        metric_out_data_path <- file.path(out_path_base,
                                          "metrics3",
                                          sprintf(out_name_base, n, p, snr_x, snr_y, loading_sparsity, rep))
        # cov_out_data_path <- file.path(out_path_base,
        #                                "covariances2",
        #                                sprintf(out_name_base, n, p, snr_x, snr_y, loading_sparsity, rep))
        # log_message("Writing metrics:", as.character(metric_out_data_path))
        
        dir.create(dirname(metric_out_data_path), recursive = TRUE, showWarnings = FALSE)
        # dir.create(dirname(cov_out_data_path), recursive = TRUE, showWarnings = FALSE)
        
        stopifnot(exists("eval_table"))
        
        saveRDS(eval_table,
                metric_out_data_path)
        
        # saveRDS(covariance_eval,
        #         cov_out_data_path)
        
        
        return(eval_table)
        
      }, error = function(e) {
        message(e)
        print("Error loading data:")
        print(model_out_data_path)
        return(NA)
      }
      )
    }
    else {return(tibble())}
    
  }
) 
# |> bind_rows()
```


```{r}

model_out_data_path = model_filenames[[1]]
print(model_out_data_path)
mod <- readRDS(model_out_data_path)

run_data <- load_data(data_path_base, data_name_base,
                              1000, 100, 1, 1, 1,
                              log = FALSE)
        
        
        jafar_posterior_means <- calc_jafar_posterior_means(mod)
        
        # Calculate structures from fitted model and simulated data
        # log_message("→ calc_all_structures est_obj")
        est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
                                       jafar_posterior_means$view_scores,
                                       jafar_posterior_means$shared_loadings,
                                       jafar_posterior_means$view_loadings)
        # log_message("→ calc_all_structures sim_obj")
        sim_obj <- calc_all_structures(run_data$Z,
                                       run_data$Phi,
                                       run_data$Lambda_l,
                                       run_data$Gamma_l)
        
        # log_message("→ eval_model + bind_rows")
        eval_table <- eval_model(est_obj,
                                 sim_obj,
                                 "covariance",
                                 "difference_norm") |> 
          mutate(n = 1000,
                 p = 100,
                 snr_x = 1,
                 snr_y = 1,
                 rep = 1
          )
        
        eval_table
```


```{r}
# getwd()
# covariance_agreement |> bind_rows() |> saveRDS("./results/jafar/covariance_agreement.rds")
```



```{r}
# covariance_agreement

jafar_cov_results |> 
  bind_rows() |> 
  mutate(n = factor(n),
         p = factor(p)) |> 
  arrange(l_1) |> 
  print() |> 
  ggplot() +
  geom_boxplot(aes(x = n, y = result, group = interaction(n, p), color = p), position = "dodge") +
  # geom_point(aes(x = n, y = result, group = p, color = p)) +
  facet_grid(l_1 ~ l_2, labeller = label_both) +
  scale_y_continuous(labels = scales::comma,
                     transform = "log10")
```

```{r}
for (snr in c(2, 1, 0.5)) {
  snr_x = snr
  snr_y = snr
  
  print(
    jafar_cov_results |> 
      mutate(n = factor(n),
             p = factor(p)) |> 
      filter(snr_x == snr,
             snr_y == snr) |> #, snr_y == 0.5) |> 
      arrange(l_1) |> 
      print() |> 
      ggplot() +
      geom_boxplot(aes(x = n, y = result, group = interaction(n, p), color = p), position = "dodge") +
      # geom_point(aes(x = n, y = result, group = p, color = p)) +
      facet_grid(l_1 ~ l_2, 
                 labeller = label_both) +
      scale_y_continuous(labels = scales::comma,
                         transform = "log10") +
      labs(
        title = paste0("Rescaled difference norm of covariances for JAFAR - SNR(X): ", snr_x, ", SNR(Y): ", snr_y),
        subtitle = TeX("$\\frac{\\| \\Lambda^{(l)}\\Lambda^{(m)T} - \\hat{\\Lambda}^{(l)}\\hat{\\Lambda}^{(m)T}\\|}{p_lp_m}$")
      )
  )
}


```


### Evaluation plots
```{r}
for (qu in c("joint_structure", "view_structure", "data_reconstruction")) {
  print(
  jafar_sim_results |> 
    mutate(n = factor(n),
           l = factor(l),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(quantity == qu) |>
    print() |>
    ggplot() +
    # geom_point(aes(x = n,
    #                y = result,
    #                # group = n,
    #                color = l)) +
    geom_boxplot(aes(x = n,
                     y = result,
                     group = interaction(n, p),
                     color = p
                     # group = n,
                     # color = l
    )) +
    facet_grid(snr_y ~ snr_x, #rows = vars(quantity),
               # cols = vars(p),
               scales = "fixed",
               labeller = label_both) + 
    labs(
      title = paste0(qu, " reconstruction relative squared error for JAFAR"),
      subtitle = TeX("$\\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$")
    )
  )
}
  
```

jafar_sim_results_withtim
```{r}
jafar_sim_results_withtime |> 
  filter(n == 500,
         p == 100,
         snr_x == 2,
         snr_y == 2,
         rep == 1)
```


```{r}
jafar_sim_results_withtime |> 
  group_by(n, p, snr_x, snr_y) |> 
  summarise(median(time)) |> 
  arrange(n, p)

jafar_sim_results_withtime |> 
  distinct(n, p, snr_x, snr_y, rep, .keep_all = TRUE) |> 
  mutate(n = factor(n),
         l = factor(l),
         p = factor(p)
         ) |>
  # distinct(n, p)
  print() |> 
  ggplot() + 
  geom_boxplot(aes(x = n, y = time)) +
  facet_wrap(~p, labeller = label_both) + 
  scale_y_continuous(trans='log10',
                     # breaks = round(seq(min(jafar_sim_results_withtime$time), max(jafar_sim_results_withtime$time), by = 200),1)
                     ) +
  labs(
    title = "Wallclock time for JAFAR runs",
    y = "Runtime seconds"
  ) + 
  guides(y = "axis_logticks")
```
# Pyro integration model loading

```{r}
pyro_metric_filenames_sparse0 <- list.files("./results/integration/sparse0/metrics/", full.names = TRUE)
pyro_metric_filenames_sparse0.25 <- list.files("./results/integration/sparse0.25/metrics/", full.names = TRUE)

pyro_sim_results_sparse0 <- lapply(pyro_metric_filenames_sparse0,
                            function(x) read_csv(x, show_col_types = FALSE)) |> 
  bind_rows() |> 
  mutate(sparsity = 0)
pyro_sim_results_sparse0.25 <- lapply(pyro_metric_filenames_sparse0.25,
                            function(x) read_csv(x, show_col_types = FALSE)) |> 
  bind_rows() |> 
  mutate(sparsity = 0.25)
```

```{r}
pyro_sim_results_sparse0.pivot <- pyro_sim_results_sparse0 |> 
  pivot_longer(cols = c('joint', 'individual', 'outcome'),
               names_to = 'structure',
               values_to = 'value') |> 
  print()

pyro_sim_results_sparse0.25.pivot <- pyro_sim_results_sparse0.25 |> 
  pivot_longer(cols = c('joint', 'individual', 'outcome'),
               names_to = 'structure',
               values_to = 'value') |> 
  print()
```

```{r}
pyro_results_total <- bind_rows(
  pyro_sim_results_sparse0.pivot,
  pyro_sim_results_sparse0.25.pivot
)
```



```{r}
pyro_sim_results.pivot |> 
  distinct(n, p, snr_x, snr_y, runtime, .keep_all = TRUE) |>
  mutate(n = factor(n),
         view = factor(view),
         p = factor(p)
         ) |>
  # distinct(n, p)
  print() |> 
  ggplot() + 
  geom_boxplot(aes(x = n, y = runtime)) +
  facet_wrap(~p, labeller = label_both) + 
  scale_y_continuous(trans='log10',
                     # breaks = round(seq(min(jafar_sim_results_withtime$time), max(jafar_sim_results_withtime$time), by = 200),1)
                     ) +
  labs(
    title = "Wallclock time for SupMultiviewDecomp runs",
    y = "Runtime seconds"
  ) + 
  guides(y = "axis_logticks")
```


```{r}
for (qu in c("joint", "individual", "outcome")) {
  print(
  pyro_sim_results.pivot |> 
    mutate(n = factor(n),
           view = factor(view),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(structure == qu,
           metric == "rse" | metric == "mse") |>
    print() |>
    ggplot() +
    # geom_point(aes(x = n,
    #                y = result,
    #                # group = n,
    #                color = l)) +
    geom_boxplot(aes(x = n,
                     y = value,
                     group = interaction(n, p),
                     color = p
                     # group = n,
                     # color = l
    )) +
    facet_grid(snr_y ~ snr_x, #rows = vars(quantity),
               # cols = vars(p),
               scales = "fixed",
               labeller = label_both) + 
    labs(
      title = paste0(qu, " reconstruction relative squared error for SupMultiviewDecomp"),
      subtitle = TeX("$\\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$")
    )
  )
}
```
```{r}
pyro_sim_results.pivot |> 
    mutate(n = factor(n),
           view = factor(view),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(structure == "individual") #,
           # metric == "95p_coverage")
```


```{r}
for (qu in c("joint", "individual", "outcome")) {
  print(
  pyro_sim_results.pivot |> 
    mutate(n = factor(n),
           view = factor(view),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(structure == qu,
           metric == "95p_coverage") |>
    print() |>
    ggplot() +
    # geom_point(aes(x = n,
    #                y = result,
    #                # group = n,
    #                color = l)) +
    geom_boxplot(aes(x = n,
                     y = value,
                     group = interaction(n, p),
                     color = p
                     # group = n,
                     # color = l
    )) +
    geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
    facet_grid(snr_y ~ snr_x, #rows = vars(quantity),
               # cols = vars(p),
               scales = "fixed",
               labeller = label_both) + 
    labs(
      title = paste0(qu, " 95% credible interval coverage probability for Pyro model"),
      subtitle = TeX("$\\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$")
    )
  )
}
```


# Combined plots

For single combination of SNR x and Y:
  - structure reconstruction for shared, individual factors
  - mse for outcome
  - outcome coverage probability
  - runtimes

Recomputing metrics for jafar does not include runtime. pull that from original metrics.
```{r}
jafar_sparse0_runtime <- jafar_sim_results_sparse0_withtime |> select(l, quantity, n, p, snr_x, snr_y, rep, time)
jafar_sparse0.25_runtime <-jafar_sim_results_sparse0.25_withtime |> select(l, quantity, n, p, snr_x, snr_y, rep, time)
```

```{r}
# Load old metric files
jafar_metrics_sparse0 <- jafar_sim_results_sparse0 |> 
  left_join(jafar_sparse0_runtime, 
            by = c('l', 'quantity', 'n', 'p', 'snr_x', 'snr_y', 'rep')) |> 
  mutate(sparsity = 0)

jafar_metrics_sparse0 <- jafar_sim_results_sparse0.25 |> 
  left_join(jafar_sparse0.25_runtime, 
            by = c('l', 'quantity', 'n', 'p', 'snr_x', 'snr_y', 'rep'))
```

```{r}
# jafar_sparse0.25_runtime
```


```{r}
jafar_sim_results
jafar_sparse0_runtime
```

```{r}
# jafar_metrics_sparse0
```


```{r}
metrics_agg
```
  
coverage of 

```{r}
jafar_metrics_sparse0.25
```

Drop rows with duplicated entries for coverage
```{r}
jafar_metrics_sparse0.25 <- metrics_agg |> 
  bind_rows() |> 
  mutate(sparsity = 0.25) |> 
  left_join(jafar_sparse0_runtime, 
            by = c('l', 'quantity', 'n', 'p', 'snr_x', 'snr_y', 'rep'))
  

jafar_coverages_sparse0.25 <- jafar_metrics_sparse0.25 |> 
  filter(metric == '95p_coverage') |> 
  distinct(l, n, p, result, .keep_all = TRUE) |> print()

jafar_metrics_sparse0.25 <- jafar_metrics_sparse0.25 |> filter(metric != '95p_coverage')

jafar_metrics_sparse0.25 <- bind_rows(jafar_metrics_sparse0.25,
                                      jafar_coverages_sparse0.25) |> 
  print()
```


```{r}
jafar_metrics_sparse0.25
```

```{r}
pyro_results_total <- bind_rows(
  pyro_sim_results_sparse0.pivot,
  pyro_sim_results_sparse0.25.pivot
) |> print()
```

```{r}
jafar_metrics_sparse0 <- jafar_metrics_sparse0 |> mutate(sparsity = 0)
jafar_metrics_sparse0.25
```

```{r}
jafar_metrics_sparse0
```

Table of aggregated metrics across all sim runs
```{r}
all_model_metrics_combined <- bind_rows(
  bind_rows(jafar_metrics_sparse0,
            jafar_metrics_sparse0.25) |> 
    mutate(model = "JAFAR",
           quantity = case_when(
             quantity == "joint_structure" ~ "joint",
             quantity == "view_structure" ~ "individual",
             .default = quantity
           )) |> 
    rename(structure = quantity,
           value = result,
           runtime = time,
           view = l),
  bind_rows(
    pyro_sim_results_sparse0.pivot,
    pyro_sim_results_sparse0.25.pivot) |> 
    mutate(model = "SupMultivewDecomp")#,
  # |> 
  #  mutate(model = "SupMultivewDecomp")
) |>
  mutate(n = factor(n),
         view = factor(view),
         p = factor(p)) |> 
  print()
```




```{r}
all_model_metrics_combined |> 
  mutate(n = factor(n),
         view = factor(view),
         p = factor(p)) 
```



```{r}
# c(0.5, 1, 2)

for (snr in c(2)) {
  for (struct in c('individual', 'joint', 'outcome')) {
  SNR_X_FILTER = snr
  SNR_Y_FILTER = snr
  
  print(all_model_metrics_combined |> 
          filter(snr_x == SNR_X_FILTER, snr_y == SNR_Y_FILTER,
                 structure == struct) |> 
                 # structure != "data_reconstruction"
          filter(metric != "95p_coverage") |> 
          ggplot() +
          geom_boxplot(aes(x = n,
                           y = value,
                           group = interaction(n, model),
                           color = model
                           # group = n,
                           # color = l
          )) + 
          facet_grid(cols = vars(p), 
                     rows = vars(sparsity), #snr_y ~ snr_x, #rows = vars(quantity),
                     # cols = vars(p),
                     scales = "fixed",
                     labeller = label_both) +
          labs(
            title = paste0("Comparison of squared errors in ", struct, " reconstruction"),
            subtitle = paste0("SNR: ", snr, " -- RSE for structures, MSE for outcome")
          )
  )
  }
}
  
  
```

```{r}
SNR_X_FILTER = 2
SNR_Y_FILTER = 2

all_model_metrics_combined |> 
  filter(metric == '95p_coverage',
         snr_x == SNR_X_FILTER, snr_y == SNR_Y_FILTER) |> 
  print() |> 
  ggplot() +
    geom_boxplot(aes(x = n,
                     y = value,
                     group = interaction(n, model),
                     color = model
                     # group = n,
                     # color = l
    )) +
    geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
    # facet_wrap(vars(p), labeller = label_both)# |> 
    facet_grid(#snr_y ~ snr_x, #
          rows = vars(sparsity),
               cols = vars(p),
               scales = "fixed",
               labeller = label_both) +
    labs(
      title = paste0("95% credible interval coverage probability for outcome")
    )
```


```{r}
all_model_metrics_combined |> 
  distinct(model, n, p, snr_x, snr_y, sparsity, runtime, .keep_all = TRUE) |>
  # mutate(n = factor(n),
  #        view = factor(view),
  #        p = factor(p)
  #        ) |>
  # distinct(n, p)
  print() |> 
  ggplot() + 
  geom_boxplot(aes(x = n, 
                   y = runtime,
                   group = interaction(n, model),
                  color = model)) +
  # facet_wrap(~p, labeller = label_both) + 
  facet_grid(cols = vars(p),
             rows = vars(sparsity),
             labeller = label_both) +
  scale_y_continuous(trans='log10',
                     # breaks = round(seq(min(jafar_sim_results_withtime$time), max(jafar_sim_results_withtime$time), by = 200),1)
                     ) +
  labs(
    title = "Wallclock time for model runs",
    y = "Runtime seconds"
  ) + 
  guides(y = "axis_logticks")
```



```{r}
tibble(pred = y_pred.means,
       outcome = y.clean,
       lower = y_lower,
       upper = y_upper,
       model = "JAFAR") |> 
  arrange(pred) |> 
  mutate(id = row_number()) |> 
  ggplot() +
  geom_errorbar(aes(x = id, ymin = lower, ymax = upper, color = model)) +
  geom_point(aes(x = id, y = pred, color = model)) +
  geom_point(aes(x = id, y = outcome)) +
  # stat_smooth(method = lm) + 
  # geom_abline(slope=1, intercept = 0) +
  labs(x = "Simulated outcome",
       y = "Predicted outcome")
```

