---
title: "Modeling"
author: "Jonathan Hori"
date: "2025-10-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
# From GitHub - requires devtools
# devtools::install_github("niccoloanceschi/jafar")


library(jafar)
library(Rfast)
library(ggplot2)
library(pROC)
library(tidyverse)
library(latex2exp)

source("../R/data_methods.R")

jafar_data <- Data

output_dir = "./results/"
dir.create(output_dir) #, showWarnings = FALSE)
# getwd()
```


```{r}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("MOFA2")

```



```{r}
library(MOFA2)
```


```{r}
getwd()
```


MCMC params
```{r}
tMCMC = 10000
tBurnIn = 5000
tThin = 10

mcmc_seed=1234
```


# Multiple datasets

Only check SNR = 2, ignore N = 1000

```{r}
TRAINING_SPLIT = TRUE
TRAINING_SIZE = 0.8
RANDOM_SEED = 123

n_array <- c(50, 100, 500)
# n_array <- c(50, 100, 500, 1000)
p_array <- c(50, 100, 1000)
snr_x_array <- c(2) #c(2, 1, 0.5)
snr_y_array <- c(2) #c(2, 1, 0.5)
# snr_x_array <- c(2, 1, 0.5)
# snr_y_array <- c(2, 1, 0.5)
reps = 10
loading_sparsity = c(0.25) #c(0, 0.25, 0.5)

sim_grid <- expand.grid(n = n_array,
            p_l = p_array,
            snr_x = snr_x_array,
            snr_y = snr_y_array,
            sparsity = loading_sparsity,
            rep = 1:reps)

data_name_base <- "sim_data_ywithview_rep%s.rds"
data_path_base <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                      "data_ortho",
                      "n%sp%s_snr%s.%s_sparse%s")

out_path_base <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                           "results/jafar/sparse_ortho") #,
                           # paste0("jafar/sparse", loading_sparsity))
out_name_base <- "n%sp%s_snr%s.%s_sparse%s_rep%s.rds"


log_dir <- "logs"

log_message <- function(...) {
  msg <- paste(Sys.time(), paste(..., collapse = " "), "\n")
  cat(msg, file = log_file, append = TRUE)
}
```


Raw code below. Attempting to parallelize below this.
```{r}
# eval_table.collected <- tibble()
# mcmc_objects <- list()
# 
# mapply(function(n, p, snr_x, snr_y) {
#   for (rep in 1:reps) {
#     # Load + process data
#     run_data <- load_data(n, p, snr_x, snr_y, rep)
#     X_l <- run_data$X_l
#     y <- run_data$y
#     X_l.clean <- standardize_views(X_l)
#     y.clean <- standardize_outcome(y)
#     
#     # use default K0. 
#     # FIT MODEL
#     tik <- Sys.time()
#     mcmc_supervised <- gibbs_jafar(X_m = X_l.clean, 
#                                    y = y.clean, 
#                                    # K0=K0, 
#                                    # K0_m=c(10, 10, 10),
#                                    tMCMC=tMCMC, 
#                                    tBurnIn=tBurnIn, 
#                                    tThin=tThin,
#                                    hyperparams = list(seed = mcmc_seed))
#     tok <- Sys.time()
#     runtime <- tok - tik
#     
#     # Average over MCMC samples
#     jafar_posterior_means <- calc_jafar_posterior_means(mcmc_supervised)
#     
#     # Calculate structures from fitted model and simulated data
#     est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
#                                    jafar_posterior_means$view_scores,
#                                    jafar_posterior_means$shared_loadings,
#                                    jafar_posterior_means$view_loadings)
#     
#     sim_obj <- calc_all_structures(run_data$Z,
#                                    run_data$Phi,
#                                    run_data$Lambda_l,
#                                    run_data$Gamma_l)
#     
#     eval_table <- bind_rows(
#       eval_model(est_obj,
#                  sim_obj,
#                  "joint_structure",
#                  "rse"),
#       eval_model(est_obj,
#                  sim_obj,
#                  "view_structure",
#                  "rse"),
#       eval_model(est_obj,
#                  sim_obj,
#                  "data_reconstruction",
#                  "rse"),
#     ) |> 
#       mutate(n = n,
#              p = p,
#              snr_x = snr_x,
#              snr_y = snr_y,
#              rep = rep,
#              time = as.numeric(difftime(tok, tik, units = "secs")))
#     
#     eval_table.collected <- bind_rows(eval_table.collected,
#                                       eval_table)
#     mcmc_objects[paste0("n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep)] <- mcmc_supervised
#   }
# },
# sim_grid$n,
# sim_grid$p_l,
# sim_grid$snr_x,
# sim_grid$snr_y)
```



```{r}

run_and_eval_model <- function(n, p, snr_x, snr_y, rep, sparsity,
                               in_path_base,
                               in_name_base,
                               out_path_base,
                               out_name_base) {
  model_out_data_path <- file.path(out_path_base,
                             "models",
                             sprintf(out_name_base,  n, p, snr_x, snr_y, sparsity, rep))
  # Skip dataset if model already run
  if (file.exists(model_out_data_path)) {NA}
  else {
  
  # Load + process data
  run_data <- load_data(in_path_base, in_name_base,
                        n, p, snr_x, snr_y, sparsity, rep)
  X_l <- run_data$X_l
  y <- run_data$y
  
  # Perform train/test split
  if (TRAINING_SPLIT) {
    set.seed(RANDOM_SEED)
    training_idx <- sample.int(n = n, size = floor(TRAINING_SIZE * n))
    
    X_l_train <- lapply(X_l, function(X) X[training_idx, ])
    X_l_test <- lapply(X_l, function(X) X[-training_idx, ])
    
    y_train <- y[training_idx]
    y_test <- y[-training_idx]
  } else {
    X_l_train <- X_l
    X_l_test <- X_l
    
    y_train <- y
    y_test <- y
  }
  
  X_l_list_column_filters <- lapply(X_l_train, function(X) zero_variance_col_filter(X))
  
  X_l_train <- lapply(1:length(X_l_train), 
                      function(l) X_l_train[[l]][, X_l_list_column_filters[[l]]])
  X_l_test <- lapply(1:length(X_l_test), 
                      function(l) X_l_test[[l]][, X_l_list_column_filters[[l]]])
  
  # Calculate mean and sd of each feature - for rescaling simulated structures
  X_l_means <- lapply(X_l_train, function(X) {
    means <- X |> colMeans()
    means
    # t(replicate(nrow(X), means))
  })
  X_l_sds <- lapply(X_l_train, function(X) {
    sds <- X |> colVars() |> sqrt()
    sds
    # diag(1 / sds)
  })
  
  y_mean <- mean(y_train)
  y_sd <- sd(y_train)
  
  log_message("Standardizing")
  # Standardize using training set means/sds
  X_l_train.clean <- standardize_views(X_l_train,
                                       list(means = X_l_means,
                                            sds = X_l_sds))
  X_l_test.clean <- standardize_views(X_l_test,
                                       list(means = X_l_means,
                                            sds = X_l_sds))
    
  y_train.clean <- (y_train - y_mean) / y_sd
  y_test.clean <- (y_test - y_mean) / y_sd

  
  # use default K0. 
  # FIT MODEL
  log_message("Fitting:")
  tik <- Sys.time()
  tryCatch({
  mcmc_supervised <- gibbs_jafar(X_m = X_l_train.clean, 
                                 y = y_train.clean, 
                                 # K0=K0, 
                                 # K0_m=c(10, 10, 10),
                                 tMCMC=tMCMC, 
                                 tBurnIn=tBurnIn, 
                                 tThin=tThin,
                                 hyperparams = list(seed = mcmc_seed))
  }, error = function(e) {
    message(e)
    log_message(e)
  })
  tok <- Sys.time()
  runtime <- tok - tik
  
  # Export model
  # model_out_data_path <- file.path(out_path_base,
  #                            "models",
  #                            sprintf(out_name_base,  n, p, snr_x, snr_y, rep))
  log_message("Writing model:", as.character(model_out_data_path))
  dir.create(dirname(model_out_data_path), recursive = TRUE, showWarnings = TRUE)
  saveRDS(mcmc_supervised,
          model_out_data_path)
  
  tryCatch({
    post_samples_y <- obtain_outcome_predictions(mcmc_supervised,
                                                 X_l_test.clean)
    
    # simulated structures need to be filtered for training data and zero 
    #     variance column filters
    training_sim_decomp <- extract_sim_decomp(run_data)
    training_sim_decomp$Z <- training_sim_decomp$Z[training_idx, ]
    training_sim_decomp$Phi <- lapply(training_sim_decomp$Phi, 
                                  function(Phi) Phi[training_idx, ])
    training_sim_decomp$Lambda_l <- lapply(1:length(training_sim_decomp$Lambda_l), 
                                  function(l) {
                                    training_sim_decomp$Lambda_l[[l]][X_l_list_column_filters[[l]], ]
                                    })
    training_sim_decomp$Gamma_l <- lapply(1:length(training_sim_decomp$Gamma_l), 
                                  function(l) {
                                    training_sim_decomp$Gamma_l[[l]][X_l_list_column_filters[[l]], ]
                                    })
    
    eval_structures <- calc_structures_with_rescaling(mcmc_supervised,
                                                      X_l_means,
                                                      X_l_sds,
                                                      post_samples_y,
                                                      training_sim_decomp)
    
    eval_table <- eval_performance(eval_structures$sim_structures,
                                   eval_structures$est_structures,
                                   eval_structures$sim_outcome,
                                   eval_structures$est_outcome,
                                   eval_structures$outcome_intervals) |> 
          mutate(n = n,
                 p = p,
                 snr_x = snr_x,
                 snr_y = snr_y,
                 rep = rep,
                 sparsity = loading_sparsity,
                 time = as.numeric(difftime(tok, tik, units = "secs"))
                 )
    
    # covariance_eval <- eval_model(est_obj,
    #                               sim_obj,
    #                               "covariance",
    #                               "difference_norm") |> 
    #   mutate(n = n,
    #          p = p,
    #          snr_x = snr_x,
    #          snr_y = snr_y,
    #          sparsity = sparsity,
    #          rep = rep
    #   )
    
    
    ### Export results
    # eval_table.collected <- bind_rows(eval_table.collected,
    #                                   eval_table)
    metric_out_data_path <- file.path(out_path_base,
                                      "metrics",
                                      sprintf(out_name_base, n, p, snr_x, snr_y, sparsity, rep))
    cov_out_data_path <- file.path(out_path_base,
                                   "covariances",
                                   sprintf(out_name_base, n, p, snr_x, snr_y, sparsity, rep))
    log_message("Writing metrics:", as.character(metric_out_data_path))
    dir.create(dirname(metric_out_data_path), recursive = TRUE, showWarnings = TRUE)
    dir.create(dirname(cov_out_data_path), recursive = TRUE, showWarnings = TRUE)
    
    stopifnot(exists("eval_table"))
    
    saveRDS(eval_table,
            metric_out_data_path)
    
    saveRDS(covariance_eval,
            cov_out_data_path)
  }, error = function(e) {
    message(e)
    log_message(e)
  })
  }
  
}
```


```{r}

# test_run_and_eval_model <- function(n, p, snr_x, snr_y, rep,
#                                     in_path_base,
#                                     in_name_base,
#                                     out_path_base,
#                                     out_name_base) {
#   log_message("Starting run")
#   load_data(in_path_base, in_name_base,
#             n, p, snr_x, snr_y, rep)
#   log_message("data loaded")
#   
#   log_message(paste0(n, p, snr_x, snr_y, rep))
#   eval_table <- tibble(c(n, p, snr_x, snr_y, rep))
#   log_message(paste0(out_path_base, "mod_n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep))
#   
#   saveRDS(eval_table, 
#           paste0(out_path_base, "/mod_n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep, ".RDS"))
# }


```

```{r}
all(is.na(NA) & is.na(NA))
```


```{r}
params <- sim_grid[1, ]
run_and_eval_model(params$n,
                   params$p,
                   params$snr_x,
                   params$snr_y,
                   params$rep,
                   params$sparsity,
                   data_path_base,
                   data_name_base,
                   out_path_base,
                   out_name_base)
```



### Fit JAFAR in parallel
```{r}
library(foreach)
library(doParallel)
```


```{r}

n_cores <- detectCores()
n_cores

cluster <- makeCluster(n_cores - 1)
registerDoParallel(cluster)

timestamp <- format(Sys.time(), "%Y%m%d_%H%M%S")
dir.create(log_dir, showWarnings = FALSE)
log_file <- file.path(log_dir, paste0("run_", timestamp, ".log"))
file.create(log_file)

# sim_grid_test <- sim_grid |> filter(snr_x == 2, #0.5, 
#                                     # (snr_y == 0.5 | snr_y == 1), 
#                                     n < 1000,
#                                     p_l < 10000)

sim_grid <- sim_grid[1, ]

tik <- Sys.time()
# Run sim
tryCatch({
  jafar_sim_results <- foreach(i = 1:nrow(sim_grid),
                               .combine = bind_rows,
                               .packages = c("tidyverse",
                                             "jafar",
                                             "Rfast")#,
                               # .export = c("load_data",
                               #             "standardize_views",
                               #             "calc_jafar_posterior_means",
                               #             "calc_all_structures",
                               #             "tMCMC",
                               #             "tBurnIn",
                               #             "tThin",
                               #             "mcmc_seed")
  ) %dopar% {
    log_message("Worker:", paste(Sys.info()[['nodename']], Sys.getpid(), sep='-'))
    params <- sim_grid[i, ]
    run_and_eval_model(params$n,
                       params$p,
                       params$snr_x,
                       params$snr_y,
                       params$rep,
                       params$sparsity,
                       data_path_base,
                       data_name_base,
                       out_path_base,
                       out_name_base)
  }
  
  message("Stopping cluster")
  stopCluster(cl = cluster)
}, error = function(e){
  message(e)
  log_message("Worker:", paste(Sys.info()[['nodename']], Sys.getpid(), sep='-'), e)
  message("Stopping cluster")
  stopCluster(cl = cluster)
}
)
tok <- Sys.time()

print(tok - tik)
```


```{r}
jafar_sim_results
```



```{r}
for (n in n_array) {
  for (p in p_array) {
    for (rep in 1:reps) {
      # Load + process data
      run_data <- load_data(n, p, snr_x, snr_y)
      X_l <- run_data$X_l
      y <- run_data$y
      X_l.clean <- standardize_views(X_l)
      y.clean <- standardize_outcome(y)
      
      # use default K0
      mcmc_supervised <- gibbs_jafar(X_m = X_l.clean, 
                                     y = y.clean, 
                                     # K0=K0, 
                                     # K0_m=c(10, 10, 10),
                                     tMCMC=tMCMC, 
                                     tBurnIn=tBurnIn, 
                                     tThin=tThin,
                                     hyperparams = list(seed = mcmc_seed))
      
      # Average over MCMC samples
      jafar_posterior_means <- calc_jafar_posterior_means(mcmc_supervised)
      
      # Calculate structures from fitted model and simulated data
      est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
                                     jafar_posterior_means$view_scores,
                                     jafar_posterior_means$shared_loadings,
                                     jafar_posterior_means$view_loadings)
      
      sim_obj <- calc_all_structures(run_data$Z,
                                     run_data$Phi,
                                     run_data$Lambda_l,
                                     run_data$Gamma_l)
      
      eval_table <- bind_rows(
        eval_model(est_obj,
                   sim_obj,
                   "joint_structure",
                   "rse"),
        eval_model(est_obj,
                   sim_obj,
                   "view_structure",
                   "rse"),
        eval_model(est_obj,
                   sim_obj,
                   "data_reconstruction",
                   "rse"),
      ) |> 
        mutate(n = n,
               p = p,
               snr_x = snr_x,
               snr_y = snr_y,
               rep = rep)
      
      eval_table.collected <- bind_rows(eval_table.collected,
                                        eval_table)
      mcmc_objects[paste0("n", n, "p", p, "snrx", snr_x, "snry", snr_y, "rep", rep)] <- mcmc_supervised
      
    }
  }
}
```


# Evaluate

Load all metric files and plot

```{r}
jafar_metric_filenames <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/metrics", 
                               pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results <- lapply(jafar_metric_filenames,
                            function(x) readRDS(x)) |> 
  bind_rows()

```

```{r}
jafar_sim_results
```




```{r}
getwd()
####### Metrics2 - reevaluating
jafar_metric_filenames_sparse0 <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0/metrics3", 
                               pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0 <- lapply(jafar_metric_filenames_sparse0,
                            function(x) readRDS(x)) |> 
  bind_rows()


jafar_metric_filenames_sparse0.25 <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0.25/metrics2", 
                               pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0.25 <- lapply(jafar_metric_filenames_sparse0.25,
                            function(x) readRDS(x)) |> 
  bind_rows()
```


```{r}
######## With runtimes
metric_filenames_sparse0_withtime <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0/metrics/", pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0_withtime <- lapply(metric_filenames_sparse0_withtime,
                            function(x) readRDS(x)) |> 
  bind_rows()


metric_filenames_sparse0.25_withtime <- list.files("/Users/jonathanhori/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/results/jafar/sparse0.25/metrics/", pattern = "\\.rds$", full.names = TRUE)

jafar_sim_results_sparse0.25_withtime <- lapply(metric_filenames_sparse0.25_withtime,
                            function(x) readRDS(x)) |> 
  bind_rows()
```


```{r}
jafar_sparse0_runtime <- jafar_sim_results_sparse0_withtime |> select(l, quantity, n, p, snr_x, snr_y, rep, time)
jafar_sparse0.25_runtime <-jafar_sim_results_sparse0.25_withtime |> select(l, quantity, n, p, snr_x, snr_y, rep, time)
```

```{r}
# Load old metric files
jafar_metrics_sparse0 <- jafar_sim_results_sparse0 |> 
  left_join(jafar_sparse0_runtime, 
            by = c('l', 'quantity', 'n', 'p', 'snr_x', 'snr_y', 'rep')) |> 
  mutate(sparsity = 0)

jafar_metrics_sparse0.25 <- jafar_sim_results_sparse0.25 |> 
  left_join(jafar_sparse0.25_runtime, 
            by = c('l', 'quantity', 'n', 'p', 'snr_x', 'snr_y', 'rep')) |> 
  mutate(sparsity = 0.25)
```


```{r}
jafar_metrics_sparse0.25
```


```{r}
getwd()
cov_filenames <- list.files("./results/jafar/covariances2/", pattern = "\\.rds$", full.names = TRUE)

jafar_cov_results <- lapply(cov_filenames,
                            function(x) readRDS(x)) |> 
  bind_rows()
```

```{r}
metric_filenames
```


### Evaluate full metrics for fitted models
```{r}
model_filenames <- list.files("./results/jafar/models/", pattern = "\\.rds$", full.names = TRUE)

# jafar_sim_models <- lapply(metric_filenames,
#                             function(x) readRDS(x)) |> 
#   bind_rows()
```

Only need to run below if metrics/covariances weren't calculated (properly) originally, or if some metrics need to be readded for fitted models.
```{r}
# i = 1
metrics_agg <- lapply(
  1:nrow(sim_grid),
  function(i) {
    # Parameters for given run
    params <- sim_grid[i, ]
    n <- params$n
    p <- params$p
    snr_x <- params$snr_x
    snr_y <- params$snr_y
    rep <- params$rep
    
    # Only evaluate parameter combinations with a fitted model
    model_out_data_path <- file.path(out_path_base,
                                     "models",
                                     sprintf(out_name_base,  n, p, snr_x, snr_y, loading_sparsity, rep))
    
    
    if (file.exists(model_out_data_path)) {
      # print(model_out_data_path)
      
      tryCatch({
        run_data <- load_data(data_path_base, data_name_base,
                              n, p, snr_x, snr_y, loading_sparsity, rep,
                              log = FALSE)
        mod <- readRDS(model_out_data_path)
        
        eval_structures <- calc_structures_with_rescaling(mod, run_data)
        eval_table <- eval_performance(eval_structures$sim_structures,
                                       eval_structures$est_structures,
                                       eval_structures$sim_outcome,
                                       eval_structures$est_outcome,
                                       eval_structures$outcome_intervals) |> 
          mutate(n = n,
                 p = p,
                 snr_x = snr_x,
                 snr_y = snr_y,
                 rep = rep,
                 sparsity = loading_sparsity
                 # time = as.numeric(difftime(tok, tik, units = "secs"))
                 )
          
        ### Export results
        # eval_table.collected <- bind_rows(eval_table.collected,
        #                                   eval_table)
        metric_out_data_path <- file.path(out_path_base,
                                          "metrics",
                                          sprintf(out_name_base, n, p, snr_x, snr_y, loading_sparsity, rep))
        # cov_out_data_path <- file.path(out_path_base,
        #                                "covariances2",
        #                                sprintf(out_name_base, n, p, snr_x, snr_y, loading_sparsity, rep))
        # log_message("Writing metrics:", as.character(metric_out_data_path))
        
        dir.create(dirname(metric_out_data_path), recursive = TRUE, showWarnings = FALSE)
        # dir.create(dirname(cov_out_data_path), recursive = TRUE, showWarnings = FALSE)
        
        stopifnot(exists("eval_table"))
        
        saveRDS(eval_table,
                metric_out_data_path)
        
        # saveRDS(covariance_eval,
        #         cov_out_data_path)
        
        
        return(eval_table)
        
      }, error = function(e) {
        message(e)
        print("Error loading data:")
        print(model_out_data_path)
        return(NA)
      }
      )
    }
    else {return(tibble())}
    
  }
) 
# |> bind_rows()
```


```{r}

model_out_data_path = model_filenames[[1]]
print(model_out_data_path)
mod <- readRDS(model_out_data_path)

run_data <- load_data(data_path_base, data_name_base,
                              1000, 100, 1, 1, 1,
                              log = FALSE)
        
        
        jafar_posterior_means <- calc_jafar_posterior_means(mod)
        
        # Calculate structures from fitted model and simulated data
        # log_message("→ calc_all_structures est_obj")
        est_obj <- calc_all_structures(jafar_posterior_means$shared_scores,
                                       jafar_posterior_means$view_scores,
                                       jafar_posterior_means$shared_loadings,
                                       jafar_posterior_means$view_loadings)
        # log_message("→ calc_all_structures sim_obj")
        sim_obj <- calc_all_structures(run_data$Z,
                                       run_data$Phi,
                                       run_data$Lambda_l,
                                       run_data$Gamma_l)
        
        # log_message("→ eval_model + bind_rows")
        eval_table <- eval_model(est_obj,
                                 sim_obj,
                                 "covariance",
                                 "difference_norm") |> 
          mutate(n = 1000,
                 p = 100,
                 snr_x = 1,
                 snr_y = 1,
                 rep = 1
          )
        
        eval_table
```


```{r}
# getwd()
# covariance_agreement |> bind_rows() |> saveRDS("./results/jafar/covariance_agreement.rds")
```



```{r}
# covariance_agreement

jafar_cov_results |> 
  bind_rows() |> 
  mutate(n = factor(n),
         p = factor(p)) |> 
  arrange(l_1) |> 
  print() |> 
  ggplot() +
  geom_boxplot(aes(x = n, y = result, group = interaction(n, p), color = p), position = "dodge") +
  # geom_point(aes(x = n, y = result, group = p, color = p)) +
  facet_grid(l_1 ~ l_2, labeller = label_both) +
  scale_y_continuous(labels = scales::comma,
                     transform = "log10")
```

```{r}
for (snr in c(2, 1, 0.5)) {
  snr_x = snr
  snr_y = snr
  
  print(
    jafar_cov_results |> 
      mutate(n = factor(n),
             p = factor(p)) |> 
      filter(snr_x == snr,
             snr_y == snr) |> #, snr_y == 0.5) |> 
      arrange(l_1) |> 
      print() |> 
      ggplot() +
      geom_boxplot(aes(x = n, y = result, group = interaction(n, p), color = p), position = "dodge") +
      # geom_point(aes(x = n, y = result, group = p, color = p)) +
      facet_grid(l_1 ~ l_2, 
                 labeller = label_both) +
      scale_y_continuous(labels = scales::comma,
                         transform = "log10") +
      labs(
        title = paste0("Rescaled difference norm of covariances for JAFAR - SNR(X): ", snr_x, ", SNR(Y): ", snr_y),
        subtitle = TeX("$\\frac{\\| \\Lambda^{(l)}\\Lambda^{(m)T} - \\hat{\\Lambda}^{(l)}\\hat{\\Lambda}^{(m)T}\\|}{p_lp_m}$")
      )
  )
}


```


### Evaluation plots
```{r}
for (qu in c("joint_structure", "view_structure", "data_reconstruction")) {
  print(
  jafar_sim_results |> 
    mutate(n = factor(n),
           l = factor(l),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(quantity == qu) |>
    print() |>
    ggplot() +
    # geom_point(aes(x = n,
    #                y = result,
    #                # group = n,
    #                color = l)) +
    geom_boxplot(aes(x = n,
                     y = result,
                     group = interaction(n, p),
                     color = p
                     # group = n,
                     # color = l
    )) +
    facet_grid(snr_y ~ snr_x, #rows = vars(quantity),
               # cols = vars(p),
               scales = "fixed",
               labeller = label_both) + 
    labs(
      title = paste0(qu, " reconstruction relative squared error for JAFAR"),
      subtitle = TeX("$\\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$")
    )
  )
}
  
```

jafar_sim_results_withtim
```{r}
jafar_sim_results_withtime |> 
  filter(n == 500,
         p == 100,
         snr_x == 2,
         snr_y == 2,
         rep == 1)
```


```{r}
jafar_sim_results_withtime |> 
  group_by(n, p, snr_x, snr_y) |> 
  summarise(median(time)) |> 
  arrange(n, p)

jafar_sim_results_withtime |> 
  distinct(n, p, snr_x, snr_y, rep, .keep_all = TRUE) |> 
  mutate(n = factor(n),
         l = factor(l),
         p = factor(p)
         ) |>
  # distinct(n, p)
  print() |> 
  ggplot() + 
  geom_boxplot(aes(x = n, y = time)) +
  facet_wrap(~p, labeller = label_both) + 
  scale_y_continuous(trans='log10',
                     # breaks = round(seq(min(jafar_sim_results_withtime$time), max(jafar_sim_results_withtime$time), by = 200),1)
                     ) +
  labs(
    title = "Wallclock time for JAFAR runs",
    y = "Runtime seconds"
  ) + 
  guides(y = "axis_logticks")
```

# Pyro results


```{r}
pyro_metric_filenames <- list.files("./results/integration/metrics/", full.names = TRUE)
pyro_sim_results <- lapply(pyro_metric_filenames,
                            function(x) read_csv(x, show_col_types = FALSE)) |> 
  bind_rows()
```


```{r}
pyro_sim_results.pivot <- pyro_sim_results |> 
  pivot_longer(cols = c('joint', 'individual', 'outcome'),
               names_to = 'structure',
               values_to = 'value') |> 
  print()
```


```{r}
# pyro_metric_filenames_sparse0 <- list.files("./results/integration/sparse0/metrics/", full.names = TRUE)
# pyro_metric_filenames_sparse0.25 <- list.files("./results/integration/sparse0.25/metrics/", full.names = TRUE)
# 
# pyro_sim_results_sparse0 <- lapply(pyro_metric_filenames_sparse0,
#                             function(x) read_csv(x, show_col_types = FALSE)) |> 
#   bind_rows() |> 
#   mutate(sparsity = 0)
# pyro_sim_results_sparse0.25 <- lapply(pyro_metric_filenames_sparse0.25,
#                             function(x) read_csv(x, show_col_types = FALSE)) |> 
#   bind_rows() |> 
#   mutate(sparsity = 0.25)
```

```{r}
# pyro_sim_results.pivot <- pyro_sim_results |> 
#   pivot_longer(cols = c('joint', 'individual', 'outcome'),
#                names_to = 'structure',
#                values_to = 'value') |> 
#   print()
# 
# pyro_sim_results_sparse0.25.pivot <- pyro_sim_results_sparse0.25 |> 
#   pivot_longer(cols = c('joint', 'individual', 'outcome'),
#                names_to = 'structure',
#                values_to = 'value') |> 
#   print()
```

```{r}
# pyro_results_total <- bind_rows(
#   pyro_sim_results_sparse0.pivot,
#   pyro_sim_results_sparse0.25.pivot
# )
```


## Pyro plots
```{r}
pyro_sim_results.pivot |> 
  distinct(n, p, snr_x, snr_y, runtime, .keep_all = TRUE) |>
  mutate(n = factor(n),
         view = factor(view),
         p = factor(p)
         ) |>
  # distinct(n, p)
  print() |> 
  ggplot() + 
  geom_boxplot(aes(x = n, y = runtime)) +
  facet_wrap(~p, labeller = label_both) + 
  scale_y_continuous(trans='log10',
                     # breaks = round(seq(min(jafar_sim_results_withtime$time), max(jafar_sim_results_withtime$time), by = 200),1)
                     ) +
  labs(
    title = "Wallclock time for SupMultiviewDecomp runs",
    y = "Runtime seconds"
  ) + 
  guides(y = "axis_logticks")
```


```{r}
for (qu in c("joint", "individual", "outcome")) {
  print(
  pyro_sim_results.pivot |> 
    mutate(n = factor(n),
           view = factor(view),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(structure == qu,
           metric == "rse" | metric == "mse") |>
    print() |>
    ggplot() +
    # geom_point(aes(x = n,
    #                y = result,
    #                # group = n,
    #                color = l)) +
    geom_boxplot(aes(x = n,
                     y = value,
                     group = interaction(n, p),
                     color = p
                     # group = n,
                     # color = l
    )) +
    facet_grid(snr_y ~ snr_x, #rows = vars(quantity),
               # cols = vars(p),
               scales = "fixed",
               labeller = label_both) + 
    labs(
      title = paste0(qu, " reconstruction relative squared error for SupMultiviewDecomp"),
      subtitle = TeX("$\\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$")
    )
  )
}
```
```{r}
pyro_sim_results.pivot |> 
    mutate(n = factor(n),
           view = factor(view),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(structure == "individual") #,
           # metric == "95p_coverage")
```


```{r}
for (qu in c("joint", "individual", "outcome")) {
  print(
  pyro_sim_results.pivot |> 
    mutate(n = factor(n),
           view = factor(view),
           p = factor(p)) |> 
    # filter(quantity == "joint_structure") |>
    # filter(quantity == "view_structure") |>
    # filter(quantity == "data_reconstruction") |>
    filter(structure == qu,
           metric == "95p_coverage") |>
    print() |>
    ggplot() +
    # geom_point(aes(x = n,
    #                y = result,
    #                # group = n,
    #                color = l)) +
    geom_boxplot(aes(x = n,
                     y = value,
                     group = interaction(n, p),
                     color = p
                     # group = n,
                     # color = l
    )) +
    geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
    facet_grid(snr_y ~ snr_x, #rows = vars(quantity),
               # cols = vars(p),
               scales = "fixed",
               labeller = label_both) + 
    labs(
      title = paste0(qu, " 95% credible interval coverage probability for Pyro model"),
      subtitle = TeX("$\\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$")
    )
  )
}
```


# Combined plots

For single combination of SNR x and Y:
  - structure reconstruction for shared, individual factors
  - mse for outcome
  - outcome coverage probability
  - runtimes

Recomputing metrics for jafar does not include runtime. pull that from original metrics.


Drop rows with duplicated entries for coverage
```{r}
jafar_metrics_sparse0.25 <- metrics_agg |> 
  bind_rows() |> 
  mutate(sparsity = 0.25) |> 
  left_join(jafar_sparse0_runtime, 
            by = c('l', 'quantity', 'n', 'p', 'snr_x', 'snr_y', 'rep'))
  

jafar_coverages_sparse0.25 <- jafar_metrics_sparse0.25 |> 
  filter(metric == '95p_coverage') |> 
  distinct(l, n, p, result, .keep_all = TRUE) |> print()

jafar_metrics_sparse0.25 <- jafar_metrics_sparse0.25 |> filter(metric != '95p_coverage')

jafar_metrics_sparse0.25 <- bind_rows(jafar_metrics_sparse0.25,
                                      jafar_coverages_sparse0.25) |> 
  print()
```


```{r}
jafar_metrics_sparse0.25
```

```{r}
pyro_results_total <- bind_rows(
  pyro_sim_results_sparse0.pivot,
  pyro_sim_results_sparse0.25.pivot
) |> print()
```

```{r}
jafar_metrics_sparse0 <- jafar_metrics_sparse0 |> mutate(sparsity = 0)
jafar_metrics_sparse0.25
```

```{r}
jafar_metrics_sparse0
```

Table of aggregated metrics across all sim runs
```{r}
all_model_metrics_combined <- bind_rows(
  jafar_sim_results |> 
  # bind_rows(jafar_metrics_sparse0,
  #           jafar_metrics_sparse0.25) |> 
    mutate(model = "JAFAR",
           quantity = case_when(
             quantity == "joint_structure" ~ "joint",
             quantity == "view_structure" ~ "individual",
             .default = quantity
           )) |> 
    rename(structure = quantity,
           value = result,
           runtime = time,
           view = l),
  bind_rows(
    pyro_sim_results.pivot) |> 
    # pyro_sim_results_sparse0.pivot,
    # pyro_sim_results_sparse0.25.pivot) |> 
    mutate(model = "SupMultivewDecomp")#,
  # |> 
  #  mutate(model = "SupMultivewDecomp")
) |>
  mutate(n = factor(n),
         view = factor(view),
         p = factor(p)) |> 
  print()
```


```{r}
all_model_metrics_combined <- all_model_metrics_combined |> 
  mutate(model = case_when(!is.na(k) ~ paste0(model, k),
                           .default = model)) |> 
  print()
```



```{r}
all_model_metrics_combined |> 
  mutate(n = factor(n),
         view = factor(view),
         p = factor(p)) 
```



```{r}
# c(0.5, 1, 2)

for (snr in c(2)) {
  for (struct in c('individual', 'joint')) {
  SNR_X_FILTER = snr
  SNR_Y_FILTER = snr
  
  print(all_model_metrics_combined |> 
          filter(snr_x == SNR_X_FILTER, snr_y == SNR_Y_FILTER,
                 structure == struct) |> 
                 # structure != "data_reconstruction"
          filter(metric != "95p_coverage") |> 
          ggplot() +
          geom_boxplot(aes(x = n,
                           y = value,
                           group = interaction(n, model),
                           color = model
                           # group = n,
                           # color = l
          )) + 
          facet_grid(cols = vars(p), 
                     rows = vars(sparsity), #snr_y ~ snr_x, #rows = vars(quantity),
                     # cols = vars(p),
                     scales = "fixed",
                     labeller = label_both) +
          labs(
            title = paste0("Comparison of squared errors in ", struct, " reconstruction (SNR ", snr, ")"),
            y = "RSE",
            subtitle = TeX("$RSE = \\frac{\\|S - \\hat{S}\\| ^{2}}{\\|S \\|^{2}}$") #paste0("SNR: ", snr, " -- Relative squared error")
          )
  )
  }
}
  
  
```


```{r}
# c(0.5, 1, 2)

for (snr in c(2)) {
  for (struct in c('outcome')) {
  SNR_X_FILTER = snr
  SNR_Y_FILTER = snr
  
  print(all_model_metrics_combined |> 
          filter(snr_x == SNR_X_FILTER, snr_y == SNR_Y_FILTER,
                 structure == struct) |> 
                 # structure != "data_reconstruction"
          filter(metric != "95p_coverage") |> 
          ggplot() +
          geom_boxplot(aes(x = n,
                           y = value,
                           group = interaction(n, model),
                           color = model
                           # group = n,
                           # color = l
          )) + 
          facet_grid(cols = vars(p), 
                     rows = vars(sparsity), #snr_y ~ snr_x, #rows = vars(quantity),
                     # cols = vars(p),
                     scales = "fixed",
                     labeller = label_both) +
          labs(
            title = paste0("Comparison of squared errors in ", struct, " reconstruction (SNR ", snr, ")"),
            subtitle = paste0("SNR: ", snr, " -- Mean squared error"),
            y = "MSE"
          ) + 
          ylim(0, 2)
  )
  }
}
  
  
```


```{r}
SNR_X_FILTER = 2
SNR_Y_FILTER = 2

all_model_metrics_combined |> 
  filter(metric == '95p_coverage',
         snr_x == SNR_X_FILTER, snr_y == SNR_Y_FILTER) |> 
  print() |> 
  ggplot() +
    geom_boxplot(aes(x = n,
                     y = value,
                     group = interaction(n, model),
                     color = model
                     # group = n,
                     # color = l
    )) +
    geom_hline(yintercept = 0.95, linetype = "dashed", color = "red") +
    # facet_wrap(vars(p), labeller = label_both)# |> 
    facet_grid(#snr_y ~ snr_x, #
          rows = vars(sparsity),
               cols = vars(p),
               scales = "fixed",
               labeller = label_both) +
    labs(
      title = paste0("95% credible interval coverage probability for outcome"),
      y = "Coverage"
    )
```


```{r}
all_model_metrics_combined |> 
  distinct(model, n, p, snr_x, snr_y, sparsity, runtime, .keep_all = TRUE) |>
  # mutate(n = factor(n),
  #        view = factor(view),
  #        p = factor(p)
  #        ) |>
  # distinct(n, p)
  print() |> 
  ggplot() + 
  geom_boxplot(aes(x = n, 
                   y = runtime,
                   group = interaction(n, model),
                  color = model)) +
  # facet_wrap(~p, labeller = label_both) + 
  facet_grid(cols = vars(p),
             rows = vars(sparsity),
             labeller = label_both) +
  scale_y_continuous(trans='log10',
                     # breaks = round(seq(min(jafar_sim_results_withtime$time), max(jafar_sim_results_withtime$time), by = 200),1)
                     ) +
  labs(
    title = "Wallclock time for model runs",
    y = "Runtime seconds"
  ) + 
  guides(y = "axis_logticks")
```



```{r}
tibble(pred = y_pred.means,
       outcome = y.clean,
       lower = y_lower,
       upper = y_upper,
       model = "JAFAR") |> 
  arrange(pred) |> 
  mutate(id = row_number()) |> 
  ggplot() +
  geom_errorbar(aes(x = id, ymin = lower, ymax = upper, color = model)) +
  geom_point(aes(x = id, y = pred, color = model)) +
  geom_point(aes(x = id, y = outcome)) +
  # stat_smooth(method = lm) + 
  # geom_abline(slope=1, intercept = 0) +
  labs(x = "Simulated outcome",
       y = "Predicted outcome")
```

## Hyperparameter tuning plots

```{r}
file_list <- list.files(
  path = "/Users/jonathanhori/multiomic_integration/sim/results/integration/hyperparam_gridsearch/metrics",
  pattern = "min_epochs20",
  full.names = TRUE # Essential for creating correct paths
)

hyper_metrics <- lapply(file_list,
                        read_csv) |> 
  bind_rows()

hyper_metrics <- read_csv("/Users/jonathanhori/multiomic_integration/sim/results/integration/hyperparam_gridsearch/metrics/run_n500p100_snr2.2_sparse0.25_deltak10_rep1_concat.csv")
```


```{r}
lrs <- hyper_metrics |> distinct(initial_lr) |> pull()
betas <- hyper_metrics |> distinct(betas) |> pull()

hyper_metrics <- hyper_metrics |> 
  pivot_longer(cols = c('joint', 'individual', 'outcome'),
               names_to = 'structure',
               values_to = 'value') |> 
  group_by(initial_lr, betas, lr_step_size, lr_decay_factor, minibatch_size) |> 
  mutate(setting = cur_group_id())

# for (beta in betas) {
print(hyper_metrics |> 
        filter(metric == "rse") |> 
        print() |> 
        ggplot() +
        geom_point(aes(x = initial_lr, y = value, shape = structure, size = minibatch_size, color = betas), position = "jitter") + 
        facet_grid(lr_step_size ~ lr_decay_factor, labeller = label_both) +
        ylim(0, 1) +
        scale_x_continuous(breaks = lrs)
)
# }
```


```{r}
hyper_metrics |> 
  filter(metric == "rse", structure != "outcome") |> 
  group_by(setting, structure) |> 
  mutate(mean_metric_by_structure = mean(value),
         agg_metric_by_structure = case_when(
           structure == "joint" ~ min(value),
           structure == "individual" ~ max(value)
         ),
         structure_mean = structure) |> 
  group_by(setting) |> 
  pivot_wider(names_from = structure_mean, 
              values_from = agg_metric_by_structure) |> 
  mutate(joint = max(joint, na.rm = TRUE),
         individual = max(individual, na.rm = TRUE),
         average_metric = mean(mean_metric_by_structure),
         metric_diff = abs(joint - individual)) |> 
  arrange(average_metric) |> # value
  filter(metric_diff < .3)
```


```{r}
hyper_metrics |> filter(setting == 61)
```


# Simple model

Evaluating optimier combinations

```{r}
eval_data <- read_csv("~/multiomic_integration/sim/results/matrixdecomp/metrics/all_metrics_concat.csv")
```


```{r matrixdecomp-load-all}
md_metric_files <- list.files(
  "~/multiomic_integration/sim/results/matrixdecomp/metrics",
  pattern = "\\.csv$",
  full.names = TRUE
)
md_metric_files <- md_metric_files[basename(md_metric_files) != "all_metrics_concat.csv"]

eval_data <- lapply(md_metric_files, read_csv, show_col_types = FALSE) |>
  bind_rows()
```


```{r}

```


```{r}
eval_data <- eval_data |> 
  arrange(outcome_mse) |> 
  mutate(n = as.factor(n),
         p = as.factor(p))
```

```{r}
eval_data |> group_by(n, p, initial_lr, lr_step_size, lr_decay_factor, betas, minibatch_size, num_particles, min_epochs, max_epochs) |> 
  summarise(mean_cov_norm = mean(cov_norm_frob),
            mean_y_mse = mean(outcome_mse),
            n()) |> 
  arrange(mean_y_mse) |> 
  filter(n == 100)
```



```{r matrixdecomp-hp-prep}
eval_data <- eval_data |>
  mutate(
    initial_lr     = factor(initial_lr),
    minibatch_size = factor(minibatch_size),
    num_particles  = factor(num_particles),
    betas          = factor(betas),
    min_epochs     = factor(min_epochs),
    max_epochs     = factor(max_epochs),
    total_runtime  = global_runtime + local_runtime
  )
```

### Structure RSE

```{r md-hp-structure-rse}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = structure_rse,
                 group = interaction(initial_lr, minibatch_size),
                 colour = minibatch_size)) +
      geom_boxplot() +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Structure RSE — betas = %s", b),
           subtitle = TeX("$RSE = \\frac{\\|S - \\hat{S}\\|^{2}}{\\|S\\|^{2}}$"),
           x = "Learning rate", y = "Structure RSE", colour = "Minibatch size")
  )
}
```

### Covariance Frobenius error

```{r md-hp-cov-frob-minibatch}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = cov_norm_frob,
                 group = interaction(initial_lr, minibatch_size),
                 colour = minibatch_size)) +
      geom_boxplot() +
      scale_y_continuous(trans = "log10") +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Covariance Frobenius error — betas = %s", b),
           x = "Learning rate", y = "Cov Frobenius norm", colour = "Minibatch size")
  )
}
```

```{r md-hp-cov-frob-particles}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = cov_norm_frob,
                 group = interaction(initial_lr, num_particles),
                 colour = num_particles)) +
      geom_boxplot() +
      scale_y_continuous(trans = "log10") +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Covariance Frobenius error — betas = %s", b),
           x = "Learning rate", y = "Cov Frobenius norm", colour = "Num particles")
  )
}
```

```{r md-hp-cov-frob-minepochs}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = cov_norm_frob,
                 group = interaction(initial_lr, min_epochs),
                 colour = min_epochs)) +
      geom_boxplot() +
      scale_y_continuous(trans = "log10") +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Covariance Frobenius error — betas = %s", b),
           x = "Learning rate", y = "Cov Frobenius norm", colour = "Min epochs")
  )
}
```

### Outcome MSE

```{r md-hp-outcome-mse-minibatch}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = outcome_mse,
                 group = interaction(initial_lr, minibatch_size),
                 colour = minibatch_size)) +
      geom_boxplot() +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Outcome MSE — betas = %s", b),
           x = "Learning rate", y = "Outcome MSE", colour = "Minibatch size")
  )
}
```

```{r md-hp-outcome-mse-particles}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = outcome_mse,
                 group = interaction(initial_lr, num_particles),
                 colour = num_particles)) +
      geom_boxplot() +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Outcome MSE — betas = %s", b),
           x = "Learning rate", y = "Outcome MSE", colour = "Num particles")
  )
}
```

```{r md-hp-outcome-mse-minepochs}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = initial_lr, y = outcome_mse,
                 group = interaction(initial_lr, min_epochs),
                 colour = min_epochs)) +
      geom_boxplot() +
      facet_grid(n + lr_decay_factor ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("Outcome MSE — betas = %s", b),
           x = "Learning rate", y = "Outcome MSE", colour = "Min epochs")
  )
}
```

### 95% Credible interval coverage

```{r md-hp-coverage}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      pivot_longer(cols = c(structure_95p_cov, outcome_95p_cov),
                   names_to = "coverage_type", values_to = "coverage") |>
      mutate(coverage_type = recode(coverage_type,
                                    structure_95p_cov = "Structure",
                                    outcome_95p_cov   = "Outcome")) |>
      ggplot(aes(x = initial_lr, y = coverage,
                 group = interaction(initial_lr, minibatch_size),
                 colour = minibatch_size)) +
      geom_boxplot() +
      geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
      facet_grid(coverage_type + n ~ p + lr_step_size, labeller = label_both) +
      labs(title = sprintf("95%% CI coverage — betas = %s", b),
           x = "Learning rate", y = "Coverage", colour = "Minibatch size")
  )
}
```

### Runtime

```{r md-hp-runtime}
for (b in levels(eval_data$betas)) {
  print(
    eval_data |> filter(betas == b) |>
      ggplot(aes(x = minibatch_size, y = total_runtime,
                 group = interaction(minibatch_size, num_particles),
                 colour = num_particles)) +
      geom_boxplot() +
      facet_grid(n + min_epochs ~ p + initial_lr, labeller = label_both) +
      scale_y_continuous(trans = "log10") +
      guides(y = "axis_logticks") +
      labs(title = sprintf("Runtime (global + local) — betas = %s", b),
           x = "Minibatch size", y = "Total runtime (s)", colour = "Num particles")
  )
}
```


# Shared Factor Model and MOFA

## Load data

```{r shared-mofa-load}
base_dir <- "./results/integration_shared"

# Auto-discover all SharedFactorModel metric top-level dirs (exclude MOFA)
shared_top_dirs <- list.dirs(base_dir, recursive = FALSE)
shared_top_dirs <- shared_top_dirs[
  grepl("^metrics", basename(shared_top_dirs)) &
  !grepl("mofa", basename(shared_top_dirs))
]

# CSVs live directly in metrics/, but in a metrics/ subfolder for dated runs
load_run_csvs <- function(top_dir, cov_only = FALSE) {
  folder_name <- basename(top_dir)
  run_label   <- if (folder_name == "metrics") "current" else sub("^metrics_", "", folder_name)
  csv_dir     <- if (folder_name == "metrics") top_dir else file.path(top_dir, "metrics")

  files <- list.files(csv_dir, pattern = "\\.csv$", full.names = TRUE)
  files <- if (cov_only) files[grepl("_cov\\.csv$", files)] else files[!grepl("_cov\\.csv$", files)]

  if (length(files) == 0) return(tibble())
  lapply(files, read_csv, show_col_types = FALSE) |> bind_rows() |> mutate(run = run_label)
}

all_shared_raw <- lapply(shared_top_dirs, load_run_csvs) |> bind_rows() 

mofa_metric_files <- list.files(
  "./results/integration_shared/metrics_mofa",
  pattern = "\\.rds$",
  full.names = TRUE
)
mofa_metric_files <- mofa_metric_files[!grepl("_cov\\.rds$", mofa_metric_files)]

mofa_metrics_raw <- lapply(mofa_metric_files, readRDS) |> bind_rows()
```

## Reshape to long format

Pivot RSE/MSE and coverage columns to a long format with `structure`, `metric`, and `value`
columns, matching the convention used in the Combined plots section.

```{r shared-mofa-reshape}
reshape_shared_metrics <- function(df, model_name) {
  rse_long <- df |>
    pivot_longer(
      cols = c(joint_rse, test_rse, outcome_mse),
      names_to  = "structure_metric",
      values_to = "value"
    ) |>
    mutate(
      structure = case_when(
        structure_metric == "joint_rse"   ~ "joint",
        structure_metric == "test_rse"    ~ "data_reconstruction",
        structure_metric == "outcome_mse" ~ "outcome"
      ),
      metric = case_when(
        structure_metric %in% c("joint_rse", "test_rse") ~ "rse",
        structure_metric == "outcome_mse"                ~ "mse"
      )
    ) |>
    select(-structure_metric)

  cov_long <- df |>
    pivot_longer(
      cols      = any_of(c("joint_95p_coverage", "outcome_95p_coverage")),
      names_to  = "structure_metric",
      values_to = "value"
    ) |>
    mutate(
      structure = case_when(
        structure_metric == "joint_95p_coverage"   ~ "joint",
        structure_metric == "outcome_95p_coverage" ~ "outcome"
      ),
      metric = "95p_coverage"
    ) |>
    select(-structure_metric)

  bind_rows(rse_long, cov_long) |>
    mutate(model = model_name)
}

all_shared_metrics <- lapply(unique(all_shared_raw$run), function(r) {
  reshape_shared_metrics(
    all_shared_raw |> filter(run == r),
    paste0("SharedFactorModel_", r)
  )
}) |> bind_rows()

mofa_metrics <- reshape_shared_metrics(mofa_metrics_raw, "MOFA")

shared_mofa_combined <- bind_rows(all_shared_metrics, mofa_metrics) |>
  mutate(
    n       = factor(n),
    p       = factor(p),
    view    = factor(view),
    k_delta = factor(k_delta),
    model   = factor(model)
  )
```

```{r}
shared_mofa_combined <- shared_mofa_combined |> filter(model != "SharedFactorModel_052826_minibatch32")
```



## Runtime

```{r shared-mofa-runtime}
for (k_val in levels(shared_mofa_combined$k_delta)) {
  print(
    shared_mofa_combined |>
      filter(k_delta == k_val) |>
      distinct(model, n, p, snr_x, snr_y, rep, sparsity, runtime) |>
      ggplot(aes(x = n, y = runtime,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      facet_wrap(~p, labeller = label_both) +
      scale_y_continuous(trans = "log10") +
      guides(y = "axis_logticks") +
      labs(title = paste0("Wallclock time — k_delta = ", k_val),
           y = "Runtime (s)")
  )
}
```

## Structure RSE

```{r shared-mofa-rse}
for (k_val in levels(shared_mofa_combined$k_delta)) {
  for (struct in c("joint", "data_reconstruction")) {
    print(
      shared_mofa_combined |>
        filter(k_delta == k_val, structure == struct, metric == "rse") |>
        ggplot(aes(x = n,
                   y = value,
                   group = interaction(n, model),
                   colour = model)) +
        geom_boxplot() +
        facet_grid(view ~ p,
                   scales = "free_y",
                   labeller = label_both) +
        labs(
          title = paste0(struct, " RSE — k_delta = ", k_val),
          subtitle = TeX("$RSE = \\frac{\\|S - \\hat{S}\\|^{2}}{\\|S\\|^{2}}$"),
          y = "RSE"
        )
    )
  }
}
```

## Outcome MSE

```{r shared-mofa-outcome-mse}
for (k_val in levels(shared_mofa_combined$k_delta)) {
  print(
    shared_mofa_combined |>
      filter(k_delta == k_val, structure == "outcome", metric == "mse", view == 0) |>
      ggplot(aes(x = n,
                 y = value,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      facet_wrap(vars(p),
                 scales = "free_y",
                 labeller = label_both) +
      labs(
        title = paste0("Outcome MSE — k_delta = ", k_val),
        y = "MSE"
      ) +
      ylim(0, 1)
  )
}
```

## 95% Credible interval coverage

```{r shared-mofa-coverage}
for (k_val in levels(shared_mofa_combined$k_delta)) {
  print(
    shared_mofa_combined |>
      filter(k_delta == k_val, metric == "95p_coverage") |>
      ggplot(aes(x = n,
                 y = value,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
      facet_grid(structure ~ p,
                 scales = "fixed",
                 labeller = label_both) +
      labs(
        title = paste0("95% CI coverage — k_delta = ", k_val),
        y = "Coverage"
      )
  )
}
```

## Covariance Frobenius error

```{r shared-mofa-cov-load}
all_shared_cov_raw <- lapply(shared_top_dirs, function(d) {
  load_run_csvs(d, cov_only = TRUE) |>
    mutate(model = paste0("SharedFactorModel_", run))
}) |> bind_rows()

mofa_cov_files <- list.files(
  "./results/integration_shared/metrics_mofa",
  pattern = "_cov\\.rds$",
  full.names = TRUE
)
mofa_cov_raw <- lapply(mofa_cov_files, readRDS) |>
  bind_rows() |>
  mutate(model = "MOFA")

shared_mofa_cov <- bind_rows(all_shared_cov_raw, mofa_cov_raw) |>
  mutate(n = factor(n), p = factor(p), k_delta = factor(k_delta), model = factor(model))
```
```{r}
shared_mofa_cov <- shared_mofa_cov |> filter(model != "SharedFactorModel_052826_minibatch32")
```


### Plot
```{r shared-mofa-cov-plot}
for (k_val in levels(shared_mofa_cov$k_delta)) {
  print(
    shared_mofa_cov |>
      filter(k_delta == k_val) |>
      ggplot(aes(x = n,
                 y = cov_frob,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      facet_grid(cov_type ~ p, labeller = label_both) +
      scale_y_continuous(trans = "log10") +
      guides(y = "axis_logticks") +
      labs(
        title = paste0("Covariance Frobenius error — k_delta = ", k_val),
        y = "Frobenius norm"
      )
  )
}
```


# Misaligned Simulations

Comparing SharedFactorModel and MOFA across four loading-alignment conditions:
`symmetric`, `mild_misalignment`, `strong_misalignment`, `anti_aligned`.

## Load data

```{r misaligned-load}
misaligned_metric_files <- list.files(
  "./results/integration_shared_misaligned/metrics",
  pattern = "[.]csv$",
  full.names = TRUE
)
misaligned_metric_files <- misaligned_metric_files[!grepl("_cov[.]csv$", misaligned_metric_files)]

misaligned_shared_raw <- lapply(misaligned_metric_files, read_csv, show_col_types = FALSE) |>
  bind_rows()

misaligned_mofa_files <- list.files(
  "./results/integration_shared_misaligned/metrics_mofa",
  pattern = "[.]rds$",
  full.names = TRUE
)
misaligned_mofa_files <- misaligned_mofa_files[!grepl("_cov[.]rds$", misaligned_mofa_files)]

misaligned_mofa_raw <- lapply(misaligned_mofa_files, readRDS) |>
  bind_rows()
```

## Reshape

```{r misaligned-reshape}
misaligned_shared_metrics <- reshape_shared_metrics(misaligned_shared_raw, "SharedFactorModel")
misaligned_mofa_metrics   <- reshape_shared_metrics(misaligned_mofa_raw,   "MOFA")

misaligned_combined <- bind_rows(misaligned_shared_metrics, misaligned_mofa_metrics) |>
  mutate(
    n           = factor(n),
    p           = factor(p),
    view        = factor(view),
    k_delta     = factor(k_delta),
    model       = factor(model),
    align_label = factor(align_label,
                         levels = c("symmetric", "mild_misalignment",
                                    "strong_misalignment", "anti_aligned"))
  )
```

```{r}
misaligned_shared_raw

misaligned_shared_metrics


misaligned_combined
```



## Runtime



```{r misaligned-runtime}
for (k_val in levels(misaligned_combined$k_delta)) {
  print(
    misaligned_combined |>
      filter(k_delta == k_val) |>
      mutate(
        svd_time = coalesce(svd_time, 0),
        runtime = runtime + svd_time) |>
      distinct(model, n, p, align_label, rep, runtime) |>
      ggplot(aes(x = n, y = runtime,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      facet_grid(align_label ~ p, labeller = label_both) +
      scale_y_continuous(trans = "log10") +
      guides(y = "axis_logticks") +
      labs(title = paste0("Wallclock time — k_delta = ", k_val),
           y = "Runtime (s)")
  )
}
```

## Structure RSE

```{r misaligned-rse}
for (k_val in levels(misaligned_combined$k_delta)) {
  for (struct in c("joint", "data_reconstruction")) {
    print(
      misaligned_combined |>
        filter(k_delta == k_val, structure == struct, metric == "rse") |>
        ggplot(aes(x = n,
                   y = value,
                   group = interaction(n, model),
                   colour = model)) +
        geom_boxplot() +
        facet_grid(align_label ~ p,
                   scales = "free_y",
                   labeller = label_both) +
        labs(
          title = paste0(struct, " RSE — k_delta = ", k_val),
          subtitle = TeX("$RSE = \\frac{\\|S - \\hat{S}\\|^{2}}{\\|S\\|^{2}}$"),
          y = "RSE"
        )
    )
  }
}
```

## Outcome MSE

```{r misaligned-outcome-mse}
for (k_val in levels(misaligned_combined$k_delta)) {
  print(
    misaligned_combined |>
      filter(k_delta == k_val, structure == "outcome", metric == "mse", view == 0) |>
      ggplot(aes(x = n,
                 y = value,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      facet_grid(align_label ~ p,
                 scales = "free_y",
                 labeller = label_both) +
      labs(
        title = paste0("Outcome MSE — k_delta = ", k_val),
        y = "MSE"
      ) +
      ylim(0, NA)
  )
}
```

```{r}
misaligned_combined |> 
  filter(model == "MOFA", metric == "mse", n == 100, p == 100)
```


## 95% Credible interval coverage

```{r misaligned-coverage}
for (k_val in levels(misaligned_combined$k_delta)) {
  print(
    misaligned_combined |>
      filter(k_delta == k_val, metric == "95p_coverage") |>
      ggplot(aes(x = n,
                 y = value,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      geom_hline(yintercept = 0.95, linetype = "dashed", colour = "red") +
      facet_grid(align_label + structure ~ p,
                 scales = "fixed",
                 labeller = label_both) +
      labs(
        title = paste0("95% CI coverage — k_delta = ", k_val),
        y = "Coverage"
      )
  )
}
```

## Covariance Frobenius error

```{r misaligned-cov-load}
misaligned_cov_files <- list.files(
  "./results/integration_shared_misaligned/metrics",
  pattern = "_cov[.]csv$",
  full.names = TRUE
)
misaligned_shared_cov <- lapply(misaligned_cov_files, read_csv, show_col_types = FALSE) |>
  bind_rows() |>
  mutate(model = "SharedFactorModel")

misaligned_mofa_cov_files <- list.files(
  "./results/integration_shared_misaligned/metrics_mofa",
  pattern = "_cov[.]rds$",
  full.names = TRUE
)
misaligned_mofa_cov <- lapply(misaligned_mofa_cov_files, readRDS) |>
  bind_rows() |>
  mutate(model = "MOFA")

misaligned_cov <- bind_rows(misaligned_shared_cov, misaligned_mofa_cov) |>
  mutate(
    n           = factor(n),
    p           = factor(p),
    k_delta     = factor(k_delta),
    model       = factor(model),
    align_label = factor(align_label,
                         levels = c("symmetric", "mild_misalignment",
                                    "strong_misalignment", "anti_aligned"))
  )
```

### Plot

```{r misaligned-cov-plot}
for (k_val in levels(misaligned_cov$k_delta)) {
  print(
    misaligned_cov |>
      filter(k_delta == k_val) |>
      ggplot(aes(x = n,
                 y = cov_frob,
                 group = interaction(n, model),
                 colour = model)) +
      geom_boxplot() +
      facet_grid(align_label + cov_type ~ p, labeller = label_both) +
      scale_y_continuous(trans = "log10") +
      guides(y = "axis_logticks") +
      labs(
        title = paste0("Covariance Frobenius error — k_delta = ", k_val),
        y = "Frobenius norm"
      )
  )
}
```
