---
title: "Data Simulation"
output: html_document
---

```{r}
library(mvtnorm)
library(tidyverse)

source("../R/sim_methods.R")
# getwd()
```


```{r}
getwd()
```

Specify number of views, subjects, data dimensions, latent factor dimensions, and SNRs.
```{r}


L = 3
N = 500
p_l = 1000 #rep(1000, L)
K = 5
K_l = 5 #rep(5, L)
snr_x = 2
snr_y = 2

loading_sparsity = 0.25

```


```{r}
sim_data_spec <- lapply(1:L,
                        function(l) list(N = N,
                                         p_l = p_l,
                                         # K = K,
                                         K_l = K_l,
                                         snr_x = snr_x,
                                         sparsity = loading_sparsity) #,
                                         # snr_y = snr_y)
)
```


# Generate data views
From latent factor scores and loadings
```{r}
set.seed(123)

sim_loadings.shared <- lapply(sim_data_spec,
                       function(spec) {
                         generate_factor_loadings(spec, 
                                                  sparsity_prob = 1 - spec$sparsity, 
                                                  K)
                       }
)

# sparsity_mask.shared <- lapply(sim_data_spec,
#                        function(spec) {
#                          generate_factor_loadings(spec, 
#                                                   sparsity_prob = 1 - spec$sparsity, 
#                                                   K)
#                        }
# )

sim_factors.shared <- generate_factor_scores(sim_data_spec[[1]], K)

sim_loadings.view_specific <- lapply(sim_data_spec,
                       function(spec) {
                         generate_factor_loadings(spec, 
                                                  sparsity_prob = 1 - spec$sparsity)
                       }
)


sim_factors.view_specific <- lapply(sim_data_spec,
                       function(spec) {
                         generate_factor_scores(spec)
                       }
)


P = projection_by_qr(sim_factors.shared)
P_perp = diag(N) - P

ortho_factors.view_specific <- lapply(sim_factors.view_specific,
                                      function(scores) {
                                        ortho_scores <- P_perp %*% scores
                                        ortho_scores_sd <- sd(ortho_scores)
                                        ortho_scores / ortho_scores_sd
                                      }
)


# sparsity_mask <- create_sparsity_mask(sparsity_prob,
#                                       nrows,
#                                       ncols)
  
```

```{r}
(generate_factor_loadings(sim_data_spec[[1]], 
                         sparsity_prob = 1 - spec$sparsity, 
                         K))

lapply(sim_data_spec,
       function(spec) print(spec))
```


```{r}
qr(rnorm(n = 5 * 1000) |>  #* 
      # rbinom(n = data_spec$K * data_spec$p_l, 
      #        size = 1, 
      #        prob = sparsity_prob) |> 
      matrix(nrow = 1000,
             ncol = 5))$rank
```


```{r}
joint_structures <- lapply(1:length(sim_data_spec), 
       function(l) {
         sim_factors.shared %*% t(sim_loadings.shared[[l]]) 
         # + 
         #   sim_factors.view_specific[[l]] %*% t(sim_loadings.view_specific[[l]]) +
         #   noise_x_views[[l]]
       }
)

view_structures <- lapply(1:length(sim_data_spec), 
       function(l) {
         # sim_factors.view_specific[[l]] %*% t(sim_loadings.view_specific[[l]])
         ortho_factors.view_specific[[l]] %*% t(sim_loadings.view_specific[[l]])
         # + 
         #   sim_factors.view_specific[[l]] %*% t(sim_loadings.view_specific[[l]]) +
         #   noise_x_views[[l]]
       }
)
```

Verify:
- rank of each joint structure (should be equal to K)

```{r}
lapply(1:L,
       function(l) qr(joint_structures[[l]])$rank)

# lapply(1:L,
#        function(l) qr(sim_loadings.shared[[l]])$rank)

```

- rank of concatenated joint structure (should also be equal to K)

```{r}
joint_concat <- do.call(cbind, joint_structures)
# scores_concat <- do.call(cbind, )

dim(joint_concat)

qr(joint_concat)$rank
```

- column space of structures is equal to scores

```{r}
# svd(sim_loadings.shared[[1]])

qr(sim_factors.shared)$rank

qr(cbind(joint_concat, sim_factors.shared))$rank
```

```{r}
# qr(t(joint_concat) %*% joint_concat)$rank
```


```{r}
# dim(t(joint_concat) %*% joint_concat)
```


```{r}
# q = qr.Q(qr(sim_factors.shared))
# projection_by_qr <- function(X) {
#   q = qr.Q(qr(X))
#   P = q %*% t(q)
# }
# projection_by_inverse <- function(X) {
#   X %*% solve(t(X) %*% X) %*% t(X)
# }
```

```{r}
P1 = projection_by_qr(joint_concat)
P1.5 = projection_by_qr(sim_factors.shared)

P2 = projection_by_inverse(joint_concat)



P1 = projection_by_qr(sim_factors.shared)
# P1.5 = projection_by_qr(sim_factors.shared)

P2 = projection_by_inverse(sim_factors.shared)
```


```{r}
P = projection_by_qr(sim_factors.shared)
P_perp = diag(N) - P
```

```{r}
P_perp %*% 
```



```{r}
all.equal(P1, P2)
```

```{r}
qr(P1)$rank
qr(P1.5)$rank
```


```{r}
qr(joint_concat)$rank
```




```{r}
# dim(P1.5)
P1 %*% sim_factors.shared
sim_factors.shared

(P1 %*% joint_concat)[1, 1]
joint_concat[1, 1]

all.equal(P1 %*% sim_factors.shared, sim_factors.shared)
all.equal(P1 %*% joint_concat, joint_concat)

all.equal(P1.5 %*% sim_factors.shared, sim_factors.shared)
all.equal(P1.5 %*% joint_concat, joint_concat)
# joint_concat
```
```{r}
P1[1, 1]
```

```{r}
P1.5[1, 1]
```


```{r}
all.equal(P1, P1.5)
```




```{r}
qr(view_structures[[1]])$rank

P = projection_by_qr(sim_factors.shared)
P_perp = diag(N) - P
ortho_view_structure <- P_perp %*% view_structures[[1]]

qr(ortho_view_structure)$rank
```

- AtJ = 0 (structures are orthogonal after projection)
```{r}
# norm(t(ortho_view_structure) %*% joint_structures[[1]])
lapply(1:L,
       function(l) norm(t(view_structures[[l]]) %*% joint_structures[[l]]))

t(ortho_view_structure) %*% sim_factors.shared
```
```{r}
lapply(1:L,
       function(l) norm(t(ortho_factors.view_specific[[l]]) %*% sim_factors.shared))
```



- After projection and rescaling, entries in view-specific matrices are normal

```{r}
mean(ortho_view_structure)
sd(ortho_view_structure)

sd(ortho_view_structure / sd(ortho_view_structure))
```



```{r}
library(rbenchmark)
benchmark("qr" = {projection_by_qr(joint_concat)},
          "inverse" = {projection_by_inverse(joint_concat)})


benchmark("qr" = {qr(joint_concat)$rank},
          "inverse" = {qr(sim_factors.shared)$rank})
```


```{r}
l = 1
q = qr.Q(qr(sim_loadings.shared[[l]])) # %*% t(sim_loadings.shared[[l]])

P = q %*% t(q)
P2 = sim_loadings.shared[[l]] %*% solve(t(sim_loadings.shared[[l]]) %*% sim_loadings.shared[[l]]) %*% t(sim_loadings.shared[[l]])
```


```{r}

```


```{r}
lapply(1:length(sim_data_spec), 
       function(l) {
         q = qr.Q(qr(sim_loadings.shared[[l]] %*% t(sim_loadings.shared[[l]])
         
         # + 
         #   sim_factors.view_specific[[l]] %*% t(sim_loadings.view_specific[[l]]) +
         #   noise_x_views[[l]]
       }
)
```




```{r}
# set.seed(123)
# sample(LETTERS, 5)
# set.seed(123)
# sample(LETTERS, 5)
```


Determine the necessary variance to obtain a desired SNR.
```{r}
view_vars <- lapply(1:L,
       function(l) {
         diag(sim_loadings.shared[[l]] %*% t(sim_loadings.shared[[l]]) + 
           sim_loadings.shared[[l]] %*% t(sim_loadings.shared[[l]])) / N
       }) #|> diag()


noise_x_views <- lapply(1:length(sim_data_spec),
                  function(l) {
                    
                    noise_x <- matrix(nrow = N, ncol = sim_data_spec[[l]]$p_l)
                    for (i in 1:N) {
                      for (j in 1:sim_data_spec[[l]]$p_l) {
                        noise_x[i, j] <- rnorm(1, sd = sqrt(view_vars[[l]][[j]] / sim_data_spec[[l]]$snr_x))
                      }
                    }
                    noise_x
                  }
)
```

Replaced sim_factors.view_specific with ortho_factors.view_specific
```{r}
sim_data.x <- lapply(1:length(sim_data_spec), 
       function(l) {
         sim_factors.shared %*% t(sim_loadings.shared[[l]]) + 
           ortho_factors.view_specific[[l]] %*% t(sim_loadings.view_specific[[l]]) +
           noise_x_views[[l]]
       }
)
```

# Generate outcome

From coefficients and factor scores

```{r}

# beta_sim <- rnorm(n = K)
# 
# outcome_var <- t(beta_sim) %*% beta_sim / K
# 
# # sim_factors.shared |> colMeans()
# 
# noise_y <- rnorm(n = N, mean = 0, sd = sqrt(outcome_var / snr_y))
# 
# sim_data.y <- sim_factors.shared %*% beta_sim + noise_y

```

OR Include view specific factors in outcome model
```{r}
beta_sim <- rnorm(n = K)

beta_l <- lapply(1:L, 
       function(l) {
         rnorm(n = sim_data_spec[[l]]$K_l)
       }
)

lp_view_specific <- lapply(1:L, 
       function(l) {
         ortho_factors.view_specific[[l]] %*% beta_l[[l]] # inserted ortho factors here
       }
)

sim_beta_cat <- c(beta_sim, beta_l |> unlist())
ncoef = K + lapply(1:L, function(l) sim_data_spec[[l]]$K_l) |> unlist() |> sum()

outcome_var <- t(sim_beta_cat) %*% sim_beta_cat / ncoef

# sim_factors.shared |> colMeans()

noise_y <- rnorm(n = N, mean = 0, sd = sqrt(outcome_var / snr_y))

sim_data.y <- sim_factors.shared %*% beta_sim + 
  Reduce("+", lp_view_specific) +
  noise_y
# sim_data.y <- sim_data.y + Reduce("+", lp_view_specific)
```


# Structure

```{r}
sim_data <- list(
  L = L,
  N = N,
  p_l = lapply(sim_data.x, ncol),
  K = K,
  K_l = c(K_l, K_l, K_l),
  X_l = sim_data.x,
  y = sim_data.y,
  Lambda_l = sim_loadings.shared,
  Gamma_l = sim_loadings.view_specific,
  Z = sim_factors.shared,
  Phi = sim_factors.view_specific,
  Phi_ortho = ortho_factors.view_specific
)


  # L = L,
  #                             N = N,
  #                             p_l = c(p_l, p_l, p_l),
  #                             K = K,
  #                             K_l = c(K_l, K_l, K_l))
```


# Save
```{r}
data_name_base <- "sim_data_n%sp%s_snr%s.%s_ywithview.rds"
# saveRDS(sim_data, sprintf(data_name_base, N, p_l, snr_x, snr_y))

saveRDS(sim_data, 
        file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                  "data",
                  paste0("test_", sprintf(data_name_base, N, p_l, snr_x, snr_y)))
)

# saveRDS(sim_data.x, "data/sim_data_x.rds")
# saveRDS(sim_data.y, "data/sim_data_y.rds")
```


# Debug: check saved data

```{r}
sim_data_multiview = readRDS('~/multiomic_integration/sim/data_ortho/train/n500p100_snr2.2_sparse0.25/sim_data_ywithview_rep1.rds')
```

```{r}

```


#--------------------
# Multiple reps


```{r}
set.seed(123)

reps = 10

n_array <- c(500) #c(50, 100, 500, 1000) #, 5000)
test_n <- 100
p_array <- c(100) #c(50, 100, 1000) #, 10000)
snr_x_array <- c(2) #c(2, 1, 0.5)
snr_y_array <- c(2) #c(2, 1, 0.5)
# snr_x_array <- c(0.5)
# snr_y_array <- c(0.5)
L = 3
K = 9
K_l = 3
loading_sparsity_array = c(.25) #c(0, 0.25, 0.5) #.25 # 25% of entries are 0
# getwd()

sim_grid <- expand.grid(n = n_array,
                        p_l = p_array,
                        snr_x = snr_x_array,
                        snr_y = snr_y_array,
                        loading_sparsity = loading_sparsity_array)

# sim_grid <- sim_grid |> filter(n == 5000, p_l == 1000, snr_x == 1, snr_y == 1)

data_name_base <- "sim_data_ywithview_rep%s.rds"
data_path_base <- "n%sp%s_snr%s.%s_sparse%s"

mapply(function(n, p_l, snr_x, snr_y, loading_sparsity) {
  # Create output path
  out_path_train <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data_ortho",
    "k.9_kl.3",
    "train",
    sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)
  )
  out_path_test <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data_ortho",
    "k.9_kl.3",
    "test",
    sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)
  )
  # if (dir.exists(out_path)) {NA}
  # else {
    dir.create(out_path_train,
               showWarnings = FALSE,
               recursive = TRUE)
    dir.create(out_path_test,
               showWarnings = FALSE,
               recursive = TRUE)
    
    # Replicates
    for (rep in 1:reps) {
      if (n == 5000 & p_l == 10000) {NA} # takes too long
      else {
        # Check if replicate already created
        if (file.exists(file.path(out_path_test,
                                  sprintf(data_name_base, rep)))) {NA}
        else {
          sim_data <- generate_factor_data(
            L = L,
            N = n,
            p_l_vec = c(p_l, p_l, p_l),
            K = K,
            K_l_vec = c(K_l, K_l, K_l),
            snr_x_vec = c(snr_x, snr_x, snr_x),
            sparsity_vec = c(loading_sparsity, loading_sparsity, loading_sparsity),
            snr_y = snr_y,
            # test_n = test_n,
            outcome = "all"
          )
          
          test_data <- generate_factor_data_testing(
            L = L,
            N = test_n,
            p_l_vec = c(p_l, p_l, p_l),
            K = K,
            K_l_vec = c(K_l, K_l, K_l),
            snr_x_vec = c(snr_x, snr_x, snr_x),
            sparsity_vec = c(loading_sparsity, loading_sparsity, loading_sparsity),
            snr_y = snr_y,
            training_loadings.shared = sim_data$Lambda_l,
            training_loadings.view_specific = sim_data$Gamma_l,
            training_beta = sim_data$beta,
            # test_n = test_n,
            outcome = "all"
          )
          
          saveRDS(sim_data, 
                  file.path(out_path_train,
                            sprintf(data_name_base, rep))
          )
          saveRDS(test_data, 
                  file.path(out_path_test,
                            sprintf(data_name_base, rep))
          )
        }
      }
    }
  # }
},
sim_grid$n,
sim_grid$p_l,
sim_grid$snr_x,
sim_grid$snr_y,
sim_grid$loading_sparsity)

```


#--------------------
# Multiple reps - single view (MatrixDecomp evaluation)

```{r}
set.seed(123)

reps      <- 10
test_n    <- 100

n_array           <- c(100, 500)
p_array           <- c(100, 500)
K_array           <- c(5, 10)
snr_x_array       <- c(1, 2)
snr_y_array       <- c(1, 2)
loading_sparsity  <- 0.25

data_name_base   <- "sim_data_rep%s.rds"
data_path_base   <- "n%sp%s_snr%s.%s_sparse%s"

sim_grid_sv <- expand.grid(
  n             = n_array,
  p_l           = p_array,
  K             = K_array,
  snr_x         = snr_x_array,
  snr_y         = snr_y_array,
  loading_sparsity = loading_sparsity,
  stringsAsFactors = FALSE
)

mapply(function(n, p_l, K, snr_x, snr_y, loading_sparsity) {
  cond_dir <- sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)
  k_dir    <- sprintf("k.%s", K)

  out_path_train <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data", "single_view", k_dir, "train", cond_dir
  )
  out_path_test <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data", "single_view", k_dir, "test", cond_dir
  )

  dir.create(out_path_train, showWarnings = FALSE, recursive = TRUE)
  dir.create(out_path_test,  showWarnings = FALSE, recursive = TRUE)

  for (rep in 1:reps) {
    train_file <- file.path(out_path_train, sprintf(data_name_base, rep))
    test_file  <- file.path(out_path_test,  sprintf(data_name_base, rep))

    if (file.exists(test_file)) next

    train_data <- generate_single_view_factor_data(
      N = n, p_l = p_l, K = K,
      snr_x = snr_x, snr_y = snr_y,
      sparsity = loading_sparsity
    )
    test_data <- generate_single_view_factor_data_testing(
      N = test_n, p_l = p_l, K = K,
      snr_x = snr_x, snr_y = snr_y,
      sparsity = loading_sparsity,
      training_loadings = train_data$Lambda,
      training_beta     = train_data$beta
    )

    saveRDS(train_data, train_file)
    saveRDS(test_data,  test_file)
  }
},
sim_grid_sv$n,
sim_grid_sv$p_l,
sim_grid_sv$K,
sim_grid_sv$snr_x,
sim_grid_sv$snr_y,
sim_grid_sv$loading_sparsity)
```


#--------------------
# Multiple reps - shared multiview (SupMultiviewShared evaluation)

```{r}
set.seed(123)

reps     <- 10
test_n   <- 100

L <- 3
K <- 5

n_array                <- c(100, 500, 1000)
p_array                <- c(50, 100, 500, 1000, 5000)
snr_x_array            <- c(2, 0.5)
snr_y_array            <- c(2, 0.5)
loading_sparsity_array <- c(0.25)

data_name_base <- "sim_data_shared_rep%s.rds"
data_path_base <- "n%sp%s_snr%s.%s_sparse%s"

sim_grid_shared <- expand.grid(
  n                = n_array,
  p_l              = p_array,
  snr_x            = snr_x_array,
  snr_y            = snr_y_array,
  loading_sparsity = loading_sparsity_array,
  stringsAsFactors = FALSE
)

mapply(function(n, p_l, snr_x, snr_y, loading_sparsity) {
  cond_dir <- sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)

  out_path_train <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data", "multi_view_shared", "k.5", "train", cond_dir
  )
  out_path_test <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data", "multi_view_shared", "k.5", "test", cond_dir
  )

  dir.create(out_path_train, showWarnings = FALSE, recursive = TRUE)
  dir.create(out_path_test,  showWarnings = FALSE, recursive = TRUE)

  for (rep in 1:reps) {
    test_file <- file.path(out_path_test, sprintf(data_name_base, rep))
    if (file.exists(test_file)) next

    train_data <- generate_factor_data_shared(
      L            = L,
      N            = n,
      p_l_vec      = rep(p_l, L),
      K            = K,
      snr_x_vec    = rep(snr_x, L),
      sparsity_vec = rep(loading_sparsity, L),
      snr_y        = snr_y
    )

    test_data <- generate_factor_data_shared_testing(
      L                 = L,
      N                 = test_n,
      p_l_vec           = rep(p_l, L),
      K                 = K,
      snr_x_vec         = rep(snr_x, L),
      sparsity_vec      = rep(loading_sparsity, L),
      snr_y             = snr_y,
      training_loadings = train_data$Lambda_l,
      training_beta     = train_data$beta
    )

    saveRDS(train_data, file.path(out_path_train, sprintf(data_name_base, rep)))
    saveRDS(test_data,  test_file)
  }
},
sim_grid_shared$n,
sim_grid_shared$p_l,
sim_grid_shared$snr_x,
sim_grid_shared$snr_y,
sim_grid_shared$loading_sparsity)
```


#--------------------
# Asymmetric variance / beta-alignment scenarios (Scenario 1)

Four alignment conditions at K=5 shared factors:
 symmetric         — all factors equally variable and equally outcome-relevant
 mild_misalignment — top-2 factors drive X but not y
 strong_misalignment — top-3 factors drive X but not y
 anti_aligned      — the single outcome-relevant factor has the lowest X-variance

factor_scale[k] multiplies loading column k across all views, controlling
factor k's variance contribution to X.  beta_weights[k] = 0 zeros out
factor k's regression coefficient, concentrating outcome relevance on the
remaining factors.

```{r}
set.seed(456)

reps   <- 10
test_n <- 100
L      <- 3
K      <- 5

# Grid over n and p; SNR and sparsity fixed for clean comparison of alignment
n_array_asym  <- c(100, 500, 1000)
p_array_asym  <- c(50, 100, 500, 1000)
snr_x_asym    <- 2.0
snr_y_asym    <- 2.0
sparsity_asym <- 0.25

alignment_scenarios <- list(
  symmetric           = list(factor_scale = c(1,   1,   1,    1,    1),
                              beta_weights = c(1,   1,   1,    1,    1)),
  mild_misalignment   = list(factor_scale = c(2,   2,   1,    1,    1),
                              beta_weights = c(0,   0,   1,    1,    1)),
  strong_misalignment = list(factor_scale = c(3,   2,   1,    0.5,  0.25),
                              beta_weights = c(0,   0,   0,    1,    1)),
  anti_aligned        = list(factor_scale = c(3,   2,   1,    0.5,  0.25),
                              beta_weights = c(0,   0,   0,    0,    1))
)

data_name_base_asym <- "sim_data_shared_rep%s.rds"

sim_grid_asym <- expand.grid(
  n           = n_array_asym,
  p_l         = p_array_asym,
  align_label = names(alignment_scenarios),
  stringsAsFactors = FALSE
)

mapply(function(n, p_l, align_label) {
  scenario     <- alignment_scenarios[[align_label]]
  factor_scale <- scenario$factor_scale
  beta_weights <- scenario$beta_weights

  cond_dir <- sprintf("n%sp%s_snr%s.%s_sparse%s_align_%s",
                      n, p_l, snr_x_asym, snr_y_asym, sparsity_asym, align_label)

  out_path_train <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data", "multi_view_shared_misaligned", "k.5", "train", cond_dir
  )
  out_path_test <- file.path(
    "/Users/jonathanhori/multiomic_integration/sim",
    "data", "multi_view_shared_misaligned", "k.5", "test", cond_dir
  )

  dir.create(out_path_train, showWarnings = FALSE, recursive = TRUE)
  dir.create(out_path_test,  showWarnings = FALSE, recursive = TRUE)

  for (rep in 1:reps) {
    test_file <- file.path(out_path_test, sprintf(data_name_base_asym, rep))
    if (file.exists(test_file)) next

    train_data <- generate_factor_data_shared_asymmetric(
      L            = L,
      N            = n,
      p_l_vec      = rep(p_l, L),
      K            = K,
      snr_x_vec    = rep(snr_x_asym, L),
      sparsity_vec = rep(sparsity_asym, L),
      snr_y        = snr_y_asym,
      factor_scale = factor_scale,
      beta_weights = beta_weights
    )

    test_data <- generate_factor_data_shared_asymmetric_testing(
      L                 = L,
      N                 = test_n,
      p_l_vec           = rep(p_l, L),
      K                 = K,
      snr_x_vec         = rep(snr_x_asym, L),
      sparsity_vec      = rep(sparsity_asym, L),
      snr_y             = snr_y_asym,
      training_loadings = train_data$Lambda_l,
      training_beta     = train_data$beta
    )

    saveRDS(train_data, file.path(out_path_train, sprintf(data_name_base_asym, rep)))
    saveRDS(test_data,  test_file)
  }
},
sim_grid_asym$n,
sim_grid_asym$p_l,
sim_grid_asym$align_label)
```


#--------------------
# Generate simple factorized matrix

X = Z %*% Lambda^T + E

```{r}
L = 1
N = 500
p_l = rep(100, L) # 1000
K = 5
K_l = rep(5, L)
snr_x = 2
snr_y = 2

loading_sparsity = 0.25

sim_data_spec <- lapply(1:L,
                        function(l) list(N = N,
                                         p_l = p_l,
                                         # K = K,
                                         K_l = K_l,
                                         snr_x = snr_x,
                                         sparsity = loading_sparsity) #,
                                         # snr_y = snr_y)
)
```

I don't remember why I wanted to calculate Z^(-1/2)
```{r}
fnMatSqrtInverse = function(mA) {
  # https://scicomp.stackexchange.com/questions/10375/efficient-computation-of-the-matrix-square-root-inverse
  ei = eigen(mA)
  d = ei$values
      d = (d+abs(d))/2
      d2 = 1/sqrt(d)
      d2[d == 0] = 0
      return(ei$vectors %*% diag(d2) %*% t(ei$vectors))
}
```


```{r}
set.seed(123)

sim_loadings.shared <- lapply(sim_data_spec,
                       function(spec) {
                         generate_factor_loadings(spec, 
                                                  sparsity_prob = 1 - spec$sparsity, 
                                                  K)
                       }
)

sim_factors.shared <- generate_factor_scores(sim_data_spec[[1]], K)


```


```{r}
factors.sqrt_inv = fnMatSqrtInverse(t(sim_factors.shared) %*% sim_factors.shared)
factors_adjusted = sim_factors.shared %*% factors.sqrt_inv

t(sim_factors.shared) %*% sim_factors.shared / N
t(factors_adjusted) %*% factors_adjusted #/ N
```


```{r}
view_vars <- lapply(1:L,
       function(l) {
         diag(sim_loadings.shared[[l]] %*% t(sim_loadings.shared[[l]]))
       }) #|> diag()


noise_x_views <- lapply(1:length(sim_data_spec),
                  function(l) {
                    
                    noise_x <- matrix(nrow = N, ncol = sim_data_spec[[l]]$p_l)
                    for (i in 1:N) {
                      for (j in 1:sim_data_spec[[l]]$p_l) {
                        noise_x[i, j] <- rnorm(1, sd = sqrt(view_vars[[l]][[j]] / sim_data_spec[[l]]$snr_x))
                      }
                    }
                    noise_x
                  }
)

# sim_data.x <- lapply(1:length(sim_data_spec), 
#                        function(l) {
#                          sim_factors.shared %*% t(sim_loadings.shared[[l]]) + 
#                            noise_x_views[[l]]
#                        }
#   )
sim_data.x <- lapply(1:length(sim_data_spec), 
                       function(l) {
                         sim_factors.shared %*% t(sim_loadings.shared[[l]]) + 
                           noise_x_views[[l]]
                         # factors_adjusted %*% t(sim_loadings.shared[[l]]) + 
                           # noise_x_views[[l]]
                       }
  )
```

```{r}
beta_sim <- rnorm(n = K_l)
    
outcome_var <- t(beta_sim) %*% beta_sim
noise_y <- rnorm(n = N, mean = 0, sd = sqrt(outcome_var / snr_y))

sim_data.y <- sim_factors.shared %*% beta_sim + noise_y
```



```{r}
F = sim_factors.shared %*% (t(sim_factors.shared) %*% sim_factors.shared / (N) + diag(ncol(sim_factors.shared)))

C = chol(t(sim_factors.shared) %*% sim_factors.shared / (N))# + diag(ncol(sim_factors.shared)))

fnMatSqrtInverse = function(mA) {
  ei = eigen(mA)
  d = ei$values
      d = (d+abs(d))/2
      d2 = 1/sqrt(d)
      d2[d == 0] = 0
      return(ei$vectors %*% diag(d2) %*% t(ei$vectors))
}


F.sqrt_inv = fnMatSqrtInverse((t(sim_factors.shared) %*% sim_factors.shared / (N)))# + diag(ncol(sim_factors.shared))))

F = sim_factors.shared %*% F.sqrt_inv

mean(F)
mean(sim_factors.shared)
sd(sim_factors.shared)

t(F) %*% F / N

(t(sim_factors.shared) %*% sim_factors.shared / (N) + diag(ncol(sim_factors.shared)))
C %*% t(C)

t(F) %*% F
```


```{r}

```

```{r}
factorQ = qr.Q(qr(sim_factors.shared))
sqrt(N) * factorQ

t(factorQ) %*% factorQ
```



```{r}
sim_data <- list(
  L = L,
  N = N,
  p_l = lapply(sim_data.x, ncol),
  K = K,
  K_l = c(K_l, K_l, K_l),
  sparsity = loading_sparsity,
  X_l = sim_data.x,
  beta = beta_sim,
  y = sim_data.y,
  Lambda_l = sim_loadings.shared,
  # Gamma_l = sim_loadings.view_specific,
  Z = sim_factors.shared
  # Phi = sim_factors.view_specific
)
```



```{r}
data_name_base <- "factor_data_n%sp%s_snr%s.%s_sparse%s.rds"
# saveRDS(sim_data, sprintf(data_name_base, N, p_l, snr_x, snr_y))
# saveRDS()
saveRDS(sim_data, #sim_data.x, 
        file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                  "data",
                  "single_view_withoutcome",
                  "k.5",
                  sprintf(data_name_base, N, p_l, snr_x, snr_y, loading_sparsity))
)
```


```{r}

svd(sim_data.x[[1]])
```


#--------------------
# Generate just one dataset
For debugging

```{r}

L = 3
n = 500
p_l = 100 #rep(1000, L)
K = 5
K_l = 5 #rep(5, L)
snr_x = 9
snr_y = 2

data_name_base <- "sim_data_ywithview_rep%s.rds"
data_path_base <- "n%sp%s_snr%s.%s"

out_path <- file.path(dirname(rstudioapi::getSourceEditorContext()$path), 
                      "data2",
                      sprintf(data_path_base, n, p_l, snr_x, snr_y))
out_path
dir.create(out_path, recursive = TRUE)


rep = 1

sim_data <- generate_factor_data(L = L,
                                 N = n,
                                 p_l_vec = c(p_l, p_l, p_l),
                                 K = K,
                                 K_l_vec = c(K_l, K_l, K_l),
                                 snr_x_vec = c(snr_x, snr_x, snr_x),
                                 sparsity_vec = c(loading_sparsity, loading_sparsity, loading_sparsity),
                                 snr_y = snr_y,
                                 outcome = "all")

saveRDS(sim_data, 
        file.path(out_path,
                  sprintf(data_name_base, rep))
)

```


Check one dataset
```{r}
test_data <- readRDS('~/Library/Mobile Documents/com~apple~CloudDocs/Projects/multiomic_integration/sim/data/n500p100_snr1.1/sim_data_ywithview_rep1.rds')

```


```{r}
test_data$X_l
```


#--------------------
# Generate Weibull survival outcomes for existing simulation data
# Reuses saved Z and beta; saves only the new outcome per replicate.

```{r}
set.seed(789)

weibull_alpha_array  <- c(1.0, 2.0)   # Weibull shape (concentration)
censoring_rate_array <- c(0.3, 0.5)   # target censoring fractions
reps <- 10

weibull_grid <- expand.grid(
  alpha          = weibull_alpha_array,
  censoring_rate = censoring_rate_array,
  stringsAsFactors = FALSE
)

data_path_base <- "n%sp%s_snr%s.%s_sparse%s"

weibull_name <- function(alpha, censoring_rate, rep)
  sprintf("weibull_a%.1f_c%.2f_rep%s.rds", alpha, censoring_rate, rep)
```


## Multiview with view-specific factors (data_ortho / k.9_kl.3)

```{r}
mv_n_array              <- c(500)
mv_p_array              <- c(100)
mv_snr_x_array          <- c(2)
mv_snr_y_array          <- c(2)
mv_loading_sparsity_array <- c(0.25)

mv_source_name <- "sim_data_ywithview_rep%s.rds"

sim_grid_mv <- expand.grid(
  n                = mv_n_array,
  p_l              = mv_p_array,
  snr_x            = mv_snr_x_array,
  snr_y            = mv_snr_y_array,
  loading_sparsity = mv_loading_sparsity_array,
  stringsAsFactors = FALSE
)

for (wi in seq_len(nrow(weibull_grid))) {
  alpha          <- weibull_grid$alpha[wi]
  censoring_rate <- weibull_grid$censoring_rate[wi]

  mapply(function(n, p_l, snr_x, snr_y, loading_sparsity) {
    cond_dir <- sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)

    for (split in c("train", "test")) {
      data_dir    <- file.path(
        "/Users/jonathanhori/multiomic_integration/sim",
        "data_ortho", "k.9_kl.3", split, cond_dir
      )
      weibull_dir <- file.path(data_dir, "weibull")
      dir.create(weibull_dir, showWarnings = FALSE, recursive = TRUE)

      for (rep in seq_len(reps)) {
        src_file <- file.path(data_dir,    sprintf(mv_source_name, rep))
        out_file <- file.path(weibull_dir, weibull_name(alpha, censoring_rate, rep))

        if (!file.exists(src_file) || file.exists(out_file)) next

        dat     <- readRDS(src_file)
        # beta is concatenated [shared (K), view-specific (L*K_l)]; extract shared portion
        beta_shared <- dat$beta[seq_len(dat$K)]
        outcome <- simulate_weibull_outcome(dat$Z, beta_shared, alpha, censoring_rate)
        saveRDS(outcome, out_file)
      }
    }
  },
  sim_grid_mv$n, sim_grid_mv$p_l, sim_grid_mv$snr_x,
  sim_grid_mv$snr_y, sim_grid_mv$loading_sparsity)
}
```


## Single view (MatrixDecomp)

```{r}
sv_n_array           <- c(100, 500, 1000, 5000)
sv_p_array           <- c(50, 100, 500, 1000, 5000)
sv_K_array           <- c(5)
sv_snr_x_array       <- c(2)
sv_snr_y_array       <- c(2)
sv_loading_sparsity  <- 0.25

sv_source_name <- "sim_data_rep%s.rds"

sim_grid_sv <- expand.grid(
  n                = sv_n_array,
  p_l              = sv_p_array,
  K                = sv_K_array,
  snr_x            = sv_snr_x_array,
  snr_y            = sv_snr_y_array,
  loading_sparsity = sv_loading_sparsity,
  stringsAsFactors = FALSE
)

for (wi in seq_len(nrow(weibull_grid))) {
  alpha          <- weibull_grid$alpha[wi]
  censoring_rate <- weibull_grid$censoring_rate[wi]

  mapply(function(n, p_l, K, snr_x, snr_y, loading_sparsity) {
    cond_dir <- sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)
    k_dir    <- sprintf("k.%s", K)

    for (split in c("train", "test")) {
      data_dir    <- file.path(
        "/Users/jonathanhori/multiomic_integration/sim",
        "data", "single_view", k_dir, split, cond_dir
      )
      weibull_dir <- file.path(data_dir, "weibull")
      dir.create(weibull_dir, showWarnings = FALSE, recursive = TRUE)

      for (rep in seq_len(reps)) {
        src_file <- file.path(data_dir,    sprintf(sv_source_name, rep))
        out_file <- file.path(weibull_dir, weibull_name(alpha, censoring_rate, rep))

        if (!file.exists(src_file) || file.exists(out_file)) next

        dat     <- readRDS(src_file)
        outcome <- simulate_weibull_outcome(dat$Z, dat$beta, alpha, censoring_rate)
        saveRDS(outcome, out_file)
      }
    }
  },
  sim_grid_sv$n, sim_grid_sv$p_l, sim_grid_sv$K,
  sim_grid_sv$snr_x, sim_grid_sv$snr_y, sim_grid_sv$loading_sparsity)
}
```


## Shared multiview (SupMultiviewShared)

```{r}
sh_n_array              <- c(100, 500, 1000, 5000)
sh_p_array              <- c(50, 100, 1000, 5000)
sh_snr_x_array          <- c(2)
sh_snr_y_array          <- c(2)
sh_loading_sparsity_array <- c(0.25)

sh_source_name <- "sim_data_shared_rep%s.rds"

sim_grid_shared <- expand.grid(
  n                = sh_n_array,
  p_l              = sh_p_array,
  snr_x            = sh_snr_x_array,
  snr_y            = sh_snr_y_array,
  loading_sparsity = sh_loading_sparsity_array,
  stringsAsFactors = FALSE
)

for (wi in seq_len(nrow(weibull_grid))) {
  alpha          <- weibull_grid$alpha[wi]
  censoring_rate <- weibull_grid$censoring_rate[wi]

  mapply(function(n, p_l, snr_x, snr_y, loading_sparsity) {
    cond_dir <- sprintf(data_path_base, n, p_l, snr_x, snr_y, loading_sparsity)

    for (split in c("train", "test")) {
      data_dir    <- file.path(
        "/Users/jonathanhori/multiomic_integration/sim",
        "data", "multi_view_shared", "k.5", split, cond_dir
      )
      weibull_dir <- file.path(data_dir, "weibull")
      dir.create(weibull_dir, showWarnings = FALSE, recursive = TRUE)

      for (rep in seq_len(reps)) {
        src_file <- file.path(data_dir,    sprintf(sh_source_name, rep))
        out_file <- file.path(weibull_dir, weibull_name(alpha, censoring_rate, rep))

        if (!file.exists(src_file) || file.exists(out_file)) next

        dat     <- readRDS(src_file)
        outcome <- simulate_weibull_outcome(dat$Z, dat$beta, alpha, censoring_rate)
        saveRDS(outcome, out_file)
      }
    }
  },
  sim_grid_shared$n, sim_grid_shared$p_l, sim_grid_shared$snr_x,
  sim_grid_shared$snr_y, sim_grid_shared$loading_sparsity)
}
```

